### 2016-02-08 paper.tex
Getting all LU algorithms in there

\documentclass[11pt, oneside]{article}
\usepackage[utf8]{inputenc}
\usepackage{bm}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{ptex2tex, minted}
\usepackage{xcolor}
\usepackage{cancel}
\usepackage{caption}
\usepackage{subfigure}
\usepackage{algorithm}
\usepackage{algpseudocode}

\subfiglabelskip=0pt

\definecolor{gray}{gray}{0.97}
\colorlet{commentcolour}{green!50!black}
\colorlet{stringcolour}{red!60!black}
\colorlet{keywordcolour}{magenta!90!black}
\colorlet{exceptioncolour}{yellow!50!red}
\colorlet{commandcolour}{blue!60!black}
\colorlet{numpycolour}{blue!60!green}
\colorlet{literatecolour}{magenta!90!black}
\colorlet{promptcolour}{green!50!black}
\colorlet{specmethodcolour}{violet}
\colorlet{indendifiercolour}{green!70!white}

%\newcommand{\codetitlestyle}[1]{\small\textit{#1}\hspace{0.1cm}}
\newcommand{\belowtitleskip}{2pt}%\smallskipamount}
%\newcommand{\captionposition}{t}

%\newcommand{\mmo}[1]{\emph{#1}}

%\renewcommand{\ttdefault}{pcr}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[top=3cm,bottom=4cm,left=3cm,right=3.2cm,asymmetric]{geometry}

\lstset{
numbers=none,
aboveskip=1ex,
belowskip=1ex,
basicstyle=\ttfamily\footnotesize,
}

\lstdefinestyle{pythonstyle}{
%%\lstset{
%%keepspaces=true,
language=python,
showtabs=true,
tab=,
tabsize=2,
basicstyle=\ttfamily\footnotesize,%\setstretch{.5},
stringstyle=\color{stringcolour},
showstringspaces=false,
alsoletter={1234567890},
otherkeywords={\ , \}, \{, \%, \&, \|},
keywordstyle=\color{keywordcolour}\bfseries,
emph={and,break,class,continue,def,yield,del,elif ,else,%
except,exec,finally,for,from,global,if,import,in,%
lambda,not,or,pass,print,raise,return,try,while,assert},
emphstyle=\color{blue}\bfseries,
emph={[2]True, False, None},
emphstyle=[2]\color{keywordcolour},
emph={[3]object,type,isinstance,copy,deepcopy,zip,enumerate,reversed,list,len,dict,tuple,xrange,append,execfile,real,imag,reduce,str,repr},
emphstyle=[3]\color{commandcolour},
emph={Exception,NameError,IndexError,SyntaxError,TypeError,ValueError,OverflowError,ZeroDivisionError},
emphstyle=\color{exceptioncolour}\bfseries,
%upquote=true,
morestring=[s]{"""}{"""},
morestring=[s]{'''}{'''},
commentstyle=\color{commentcolour}\slshape,
emph={[4]1, 2, 3, 4, 5, 6, 7, 8, 9, 0,ode, fsolve, sqrt, exp, sin, cos, arccos, pi, array, norm, solve,float,complex, dot, arange, isscalar, max, sum, flatten, shape, reshape, find, any, all, abs, plot, linspace, legend, quad, polyval,polyfit, hstack,vector, concatenate,vstack,column_stack,empty,zeros,ones,rand,vander,grid,pcolor,eig,eigs,eigvals,svd,qr,tan,det,logspace,roll,min,mean,cumsum,cumprod,diff,vectorize,lstsq,cla,eye,xlabel,ylabel,squeeze},
emphstyle=[4]\color{commandcolour},
emph={[5]__init__,__add__,__mul__,__div__,__sub__,__call__,__getitem__,__setitem__,__eq__,__ne__,__nonzero__,__rmul__,__radd__,__repr__,__str__,__get__,__truediv__,__pow__,__name__,__future__,__all__},
emphstyle=[5]\color{specmethodcolour},
emph={[6]assert,range,yield},
emphstyle=[6]\color{keywordcolour}\bfseries,
emph={[7]def, return, and, print},
emphstyle=[7]\color{black}\bfseries,
% emph={[7]self},
% emphstyle=[7]\bfseries,
literate=*%
%{:}{{\literatecolour:}}{1}%
%{=}{{\literatecolour=}}{1}%
%{-}{{\literatecolour-}}{1}%
%{+}{{\literatecolour+}}{1}%
%{*}{{\literatecolour*}}{1}%
{/}{{\literatecolour/}}{1}%
{!}{{\literatecolour!}}{1}%
%{(}{{\literatecolour(}}{1}%
%{)}{{\literatecolour)}}{1}%
%{[}{{\literatecolour[}}{1}%
%{]}{{\literatecolour]}}{1}%
{<}{{\literatecolour<}}{1}%
{>}{{\literatecolour>}}{1}%
{>>>}{{\textcolor{promptcolour}{>>>}}}{1}%
,%
breaklines=true,
breakatwhitespace= true,
xleftmargin=\framemargin,
xrightmargin=\framemargin,
aboveskip=1ex,
belowskip=1ex,
frame=trbl, %trbl
numbers=none,
%frameround=tttt,
rulecolor=\color{black!40},
%framexleftmargin=\framemargin,
%framextopmargin=.1ex,
%framexbottommargin=.1ex,
%framexrightmargin=\framemargin,
%framexleftmargin=1mm, framextopmargin=1mm, frame=shadowbox, rulesepcolor=\color{blue},#1
%frame=tb,
%backgroundcolor=\color{yellow!10}
backgroundcolor=\color{gray}
%}
}

\newcommand{\N}[1]{\check{#1}}
\newcommand{\D}[1]{\overline{#1}}

\newcommand{\inpyth}{\lstinline[style=pythonstyle, basicstyle=\ttfamily]} %[]%

\lstnewenvironment{inpython}[1][]{
\lstset{style=pythonstyle, frame=trbl, belowcaptionskip=\belowtitleskip}
}{}

\newcommand{\includecode}[2][py]{\lstinputlisting[caption=#2,label=list:#2,style=pythonstyle,
float=!htpb]{#2}}

\newcommand{\mmo}[1]{({\bf mmo comment:} \emph{#1})}
\bibliographystyle{plain}

\title{Assessment of a Shen-Fourier turbulent channel flow solver for large-scale simulations}
\author{Mikael Mortensen and Diako Darian}
%\date{}							% Activate to display a given date or no date

\begin{document}
\maketitle
\section*{Abstract}

\section{Introduction}

\section{Preliminaries}
The Navier-Stokes equations used to describe turbulent flow in a doubly 
periodic channel can be written in rotational form as
\begin{align}
 \frac{\partial \bm{u}}{\partial t}   &= \bm{\mathcal{H}} + \nu 
 \nabla^2 \bm{u} - \nabla{P}, \notag \\
 \nabla \cdot \bm{u} &= 0, \label{eq:NS}
\end{align}
where $\bm{u}(\bm{x}, t)=(u, v, w)$ is the velocity vector, $\bm{x}=(x, y, z)$ 
and $t$ are position and time and $\bm{\mathcal{H}}(\bm{x}, t) = \bm{u}\times 
\bm{\omega}$ (where $\bm{\omega} = \nabla \times \bm{u}$)  represents the 
nonlinear convection.  The constant dynamic viscosity is denoted as $\nu$ and 
$P$ is a pressure modified to account for the kinetic energy, i.e., $P = p + 
\bm{u} \cdot \bm{u}/2$, where $p$ the instantaneous pressure 
normalized by a constant density. The computational domain is $\Omega=[-1, 
1]\times [0, L_y] \times [0, L_z]$ with channel walls located at $x=\pm 1$ such 
that no-slip applies at $ \bm{u}(\pm 1, y, z, t) = 0$. The walls are spanning 
the $y-z$ plane and the equations are periodic in the $y$ and $z$ directions 
with periodic lengths $L_y$ and $L_z$, respectively.

The Navier Stokes equations are discretized using Fourier basis functions for the periodic directions, and a combination of Chebyshev polynomials in the wall normal direction. Three different sets of basis functions and function spaces are relevant for the wall normal direction
\begin{align}
&  \phi_k(x) = T_k(x), & V_{N_x} &= \text{span}\{\phi_0, \phi_1,\ldots, \phi_{N_x}\} \label{eq:Tk}\\
& \D{\phi}_k(x) = T_k(x) - T_{k+2}(x), & \D{V}_{N_x} &= \text{span} \{ \D{\phi}_0, \D{\phi}_1, \ldots, \D{\phi}_{N_x-2} \} \label{eq:phiD}\\
& \N{\phi}_k(x) = T_k(x) - \frac{2(k+2)}{k+3} T_{k+2}(x) + 
\frac{k+1}{k+3} T_{k+4}, & \N{V}_{N_x} &= \text{span} \{\N{\phi}_0, \N{\phi}_1, 
\ldots, \N{\phi}_{N_x-4} \} \label{eq:phiN} 
\end{align}
where $T_k(x)$ is the $k$'th degree Chebyshev polynomial of the first kind. The 
basis functions and function spaces in (\ref{eq:phiD}) and (\ref{eq:phiN}) were 
suggested by Shen, and satisfy, respectively, the boundary conditions 
$\D{\phi}_k(\pm 1) = 0$, $\N{\phi}_k(\pm 1)=0$ and $\N{\phi}'_k(\pm 1)=0$. The 
last two function spaces may alternatively be written as $\D{V}_{N_x} = \{v \in 
V_{N_x}: v(\pm 1)=0 \}$ and $\N{V}_{N_x} = \{v \in V_{N_x}: v(\pm 1) = v'(\pm 
1) = 0 \}$.

Three-dimensional basis functions and function spaces that are periodic in $y$ 
and $z$ directions, can now be defined as
\begin{align}
  \psi_{\bm{k}}(\bm{x}) = \phi_{l}(x)e^{ \imath(\underline{m} y + \underline{n} z)}, \quad V_N &= \text{span} \{ \psi_{\bm{k}}(\bm{x}):\, \bm{k} \in K_N  \}, \\
  \D{\psi}_{\bm{k}}(\bm{x}) = \D{\phi}_{l}(x)e^{ \imath(\underline{m} y + \underline{n} z)}, \quad \D{V}_N &= \text{span} \{ \D{\psi}_{\bm{k}}(\bm{x}):\, \bm{k} \in \D{K}_N  \}, \\
  \N{\psi}_{\bm{k}}(\bm{x}) = \N{\phi}_{l}(x)e^{ \imath(\underline{m} y + \underline{n} z)}, \quad \N{V}_N &= \text{span} \{ \N{\psi}_{\bm{k}}(\bm{x}):\, \bm{k} \in \N{K}_N  \},
\end{align}
where $\imath=\sqrt{-1}$ and 
\begin{align}
K_N = \Big\{\bm{k} \in \mathcal{Z}\times\mathcal{R}^2 / &(l, \underline{m}, \underline{n}) \in \left( l, \frac{2 \pi m}{L_y}, \frac{2 \pi n}{L_z} \right), \text{where} \notag \\
 &(l, m, n) \in  [0, \ldots, N_x] \times [-\frac{N_y}{2},\ldots, \frac{N_y}{2}-1] \times[-\frac{N_z}{2},\ldots, \frac{N_z}{2}-1] \Big\}. \label{eq:wavenumbermesh}
\end{align}
$\D{K}_N$ and $\N{K}_N$ are both equal to $K_N$ except from index $l$ that 
ranges $[0, 1, \ldots, N_x-2]$ and $[0, 1, \ldots, N_x-4]$ for $\D{K}_N$ 
and $\N{K}_N$ respectively, see (\ref{eq:phiD}) and (\ref{eq:phiN}). We use 
notation $K_N^p(m, n)=K_N(0, m, n)$ when referring to only the two periodic 
dimensions of $K_N$.

The computational mesh has in real space been created using $N = (N_x, N_y, 
N_z)$ 
intervals, where the two periodic directions use uniform intervals. The 
computational mesh is given as
\begin{align}
  X_N = \Big\{ \bm{x} \in \mathcal{R}^3/&(x_i, y_j, z_k) \in \left( h(i), \frac{jL_y}{N_y}, \frac{kL_z}{N_z} \right), \text{where} \notag \\
  &(i, j, k) \in [0, 1, \ldots, N_x] \times [0, 1, \ldots, N_y-1] \times [0, 1, \ldots, N_z-1] \Big\} \label{eq:Xn}
\end{align}
where $x_i = h(i)$ represents either
\begin{equation}
 h(i) = \begin{cases}
   \cos \left(\frac{i \pi }{N_x} \right) \, &\forall \, i=0,1, \ldots, N_x \quad  \text{for Gauss-Lobatto points}, \\
   \cos \left(\frac{(2i +1)\pi}{2N_x+2} \right) \, &\forall \, i=0,1, \ldots, N_x \quad  \text{for Gauss-Chebyshev points}. \\
 \end{cases}
\end{equation}

For the Navier Stokes equations we look for solutions of the velocity 
components of the form
\begin{align}
u(\bm{x}, t) &= \frac{1}{N_yN_z}\sum_{\bm{k} \in \N{K}_N} \hat{u}_{\bm{k}}(t) 
\N{\psi}_{\bm{k}}(\bm{x}), \label{eq:u_solx} \\
v(\bm{x}, t) &= \frac{1}{N_yN_z}\sum_{\bm{k} \in \D{K}_N} \hat{v}_{\bm{k}}(t) 
\D{\psi}_{\bm{k}}(\bm{x}), \label{eq:u_soly} \\
w(\bm{x}, t) &= \frac{1}{N_yN_z}\sum_{\bm{k} \in \D{K}_N} \hat{w}_{\bm{k}}(t) 
\D{\psi}_{\bm{k}}(\bm{x}), \label{eq:u_solz}
\end{align}
where $\hat{u}_{\bm{k}}(t) = \hat{u}(l, {m}, {n}, t)$ are the expansion 
coefficients for the velocity component in $x$-direction (and similar for the 
other two components) and the scaling by $N_y$ and $N_z$ is merely for 
convenience 
and compliance with the definition used for the inverse discrete Fourier 
transform. Note that from now on we will simply use the notation $\hat{u}$ for 
$\hat{u}_{\bm{k}}(t)$, when it is possible to simplify without loss of clarity. 
Likewise we will simply use $u$ for $u(\bm{x}, t)$.

If the expansion coefficients $\hat{u}$  are known for the entire wavenumber 
mesh (\ref{eq:wavenumbermesh}), then the expression (\ref{eq:u_solx}) may be 
evaluated on the mesh (\ref{eq:Xn}) using fast Fourier and fast Shen inverse 
transforms as
\begin{align}
u(x_i, y_j, z_k, t) &= \underbrace{\frac{1}{N_z}\sum_n 
\underbrace{\frac{1}{N_y} \sum_m \underbrace{\sum_l 
\hat{u}(l,m,n)\N{\phi}_l(x_i)}_{\N{\mathcal{S}}_x^{-1}} e^{\imath \underline{m} 
y_j}}_{\mathcal{F}_y^{-1}} e^{\imath \underline{n} z_k}}_{\mathcal{F}_z^{-1}} , 
\quad \forall \, \bm{x} \in X_N, \notag \\
  u &= \N{\mathcal{T}}^{-1}(\hat{u}) =  \mathcal{F}^{-1}_z (\mathcal{F}^{-1}_y 
  (\N{\mathcal{S}}^{-1}_x (\hat{u}))), \quad \forall \, \bm{x} \in X_N.  
  \label{eq:ifft} 
\end{align}
Here $\N{\mathcal{T}}^{-1}$ is used as short notation for the complete inverse 
transform, $\mathcal{F}_{y}^{-1}$ and $\mathcal{F}_{z}^{-1}$ represents inverse 
Fourier transforms along directions $y$ and $z$ respectively, and the inverse 
Shen transform $\N{\mathcal{S}}_{x}^{-1}$ is performed along the wall normal 
$x$ direction. Note that the transforms are applied sequentially to the entire 
computational mesh. The inverse Shen transform may be computed using fast 
Chebyshev transforms for all the three bases in (\ref{eq:Tk}, \ref{eq:phiD}, 
\ref{eq:phiN}). The transforms are slightly different for the three bases and 
${\mathcal{S}}, \D{\mathcal{S}}$ and $\N{\mathcal{S}}$ are used to distinguish 
between them with obvious notation.

A three-dimensional scalar product in the weighted $[L^2_w(\Omega)]^3$ space is defined as
\begin{align}
 \left<u, v\right>_w &= \int_{\Omega} {u(\bm{x}) v^*(\bm{x})}w\,dxdydz, \notag \\
   &= \int_{\Omega} {u(\bm{x}) v^*(\bm{x})}\,dx_wdydz,
\end{align}
where $v^*$ is the comlex conjugate of $v$ and the weights $w$ are unity for periodic directions and  $w=1/\sqrt{1-x^2}$ for the inhomogeneous direction, and we have used $dx_w = dx/\sqrt{1-x^2}$. In this work we will make use of the discrete weighted $[l^2_w(\Omega)]^3$ space, where quadrature is employed for integration. As such the scalar product corresponds to sums that may be computed using fast transforms
\begin{align}
 \left< u, \N{\psi}_{\bm{k}} \right>_w &= h \underbrace{\sum_i 
 \underbrace{\sum_j \underbrace{\sum_k u(x_i, y_j, z_k, t)  e^{-\imath 
 \underline{n} z_k}}_{\mathcal{F}_n}  e^{-\imath \underline{m} y_j} 
 }_{\mathcal{F}_m} \N{\phi}_{l}(x_i) w_i}_{\N{\mathcal{S}}_l},   \notag \\
  &=  h \N{\mathcal{S}}(u) = h \N{\mathcal{S}}_{l} (\mathcal{F}_{{m}} 
  (\mathcal{F}_{{n}}(u))), \quad \forall \, \bm{k} \in \N{K}_N. \label{eq:sst1}
\end{align}
Here $h\N{\mathcal{S}}$ denotes the complete three dimensional scalar product 
and $h = L_yL_zN_y^{-1}N_z^{-1}$ is a constant. $\mathcal{F}_{{n}}$ and 
$\mathcal{F}_{{m}}$ represent discrete Fourier transforms in $z$- and 
$y$-directions, respectively, and $\N{\mathcal{S}}_{l}(\cdot) = (\cdot, 
\N{\phi}_l)_w$ is the Shen scalar product in the $x$-direction. The weights, 
$w_i$, required for the Shen transforms are 
defined as
\begin{equation}
 w_i = \begin{cases}
       \frac{\pi}{c_i N_x} &\forall \, i=0,1,\ldots, N_x \quad  \text{for 
       Gauss-Lobatto points},\\
       \frac{\pi}{N_x+1} &\forall \, i=0,1,\ldots, N_x  \quad \text{for 
       Gauss-Chebyshev points},      
 \end{cases}
\end{equation}
where $c_0 = c_{N_x} = 2$ and $c_i = 1$ for $0 < i < N_x$.

The scalar product  $h\N{\mathcal{S}}$ in (\ref{eq:sst1}) does not represent a 
complete transform. To find a transformation between physical $u$ and spectral 
$\hat{u}$ we make use of Eq. (\ref{eq:u_solx}) directly and use the discrete 
orthogonality of the Fourier basis functions
\begin{align}
\left<u, \N{\psi}_{\bm{k}}\right>_w &= \frac{1}{N_yN_z}\left< 
\sum_{\bm{m}=(q,r,s) \in \N{K}_N} \hat{u}_{\bm{m}} \N{\psi}_{\bm{m}}, 
\N{\psi}_{\bm{k}} \right>_w, \notag \\
           &= h \sum_{q=0}^{N_x-4} \sum_{i=0}^{N_x} \N{\phi}_{q}(x_i) 
           \N{\phi}_{l}(x_i) w_i \, \hat{u}(q, {m}, {n}, t), \notag \\
           &= h \sum_{p=0}^{N_x-4} \N{B}_{lq} \hat{u}(q, {m}, {n}, t), 
           \label{eq:sst2}
\end{align}
where $\N{B}_{lq} = (\N{\phi}_q, \N{\phi}_l)_w = 
\sum_{i=0}^{N_x} \N{\phi}_{q}(x_i) \N{\phi}_{l}(x_i) w_i$ is a 
pentadiagonal mass matrix. The complete transformation is now obtained by 
setting Eq. (\ref{eq:sst2}) equal to (\ref{eq:sst1}) and solving for $\hat{u}$. 
Moving to matrix notation we get
\begin{align}
h\sum_{q=0}^{N_x-4} \N{B}_{lq}\, \hat{u}(q, {m}, {n}, t) &= 
h\N{\mathcal{S}}(u)(l, m, n, t), \notag \\
 \hat{u}_{\bm{k}}(t) &= \N{\mathcal{T}}(u) =  \N{B}^{-1}\N{\mathcal{S}}(u), 
 \quad \forall \, \bm{k} \in \N{K}_N. \label{eq:fstB}
\end{align}
where $\N{\mathcal{T}}(u)$ denotes the complete transformation, such that $u = 
\N{\mathcal{T}}^{-1}(\N{\mathcal{T}}(u))$. Note that since $\N{B}$ assembles to 
a pentadiagonal matrix the solution ($\N{B}^{-1}$) can be obtained very fast 
and 
the complete 
transformation thus requires a fast Chebyshev transform ($O(N_x \log N_x)$) and 
a fast linear algebraic solve ($O(N_x)$). Similar transforms $\mathcal{T}$ 
and $\D{\mathcal{T}}$ are defined for the two other bases (\ref{eq:Tk}) and 
(\ref{eq:phiD}), using mass matrices $B_{lq}=(T_q, T_l)_w$ (diagonal) and 
$\D{B}_{lq}=(\D{\phi}_l, \D{\phi}_q)_w$ (tridiagonal) and scalar products 
$\mathcal{S}_l(\cdot) = (\cdot, T_l)_w$ and $\D{\mathcal{S}}_l(\cdot) = (\cdot, 
\D{\phi}_l)_w$.

\section{Discretization of Navier Stokes equations}
\label{sec:discretizationNS}
The Navier Stokes equations (\ref{eq:NS}) are solved using a scheme 
proposed by Kim, Moin and Moser (1987). This scheme is developed by taking the 
Laplacian of the wall normal momentum equation and the curl of the momentum 
equation. Following elimination of the pressure, the equations to solve are 
\begin{align}
\frac{\partial}{\partial t} \nabla^2 u &= h_u + \nu \nabla^4 u, 
\label{eq:biharmonic} \\
\frac{\partial g}{\partial t} &= h_g + \nu \nabla^2 g, \label{eq:g} \\
f + \frac{\partial u}{\partial x} &= 0, \label{eq:f}
\end{align}
where
\begin{align}
f &= \frac{\partial v}{\partial y} + \frac{\partial w}{\partial z}, \\
g &= \frac{\partial w}{\partial y} - \frac{\partial v}{\partial z}, \\
h_u = -\frac{\partial}{\partial x} &\left( \frac{\partial 
\mathcal{H}_y}{\partial y} + \frac{\partial \mathcal{H}_z}{\partial z} \right) 
+ \left(\frac{\partial^2}{\partial y^2} + \frac{\partial^2}{\partial z^2} 
\right) \mathcal{H}_x ,
\\
h_g &= \frac{\partial \mathcal{H}_z}{\partial y} - \frac{\partial 
\mathcal{H}_y}{\partial z}.
\end{align}
The two remaining velocity components are computed from the definitions of $f$ 
and $g$. The biharmonic equation (\ref{eq:biharmonic}) is solved with boundary 
conditions are $u(\pm 1) = u'(\pm 1) = 0$, where the Neumann conditions follow 
from the continuity equation. The boundary conditions for Eq. (\ref{eq:g}) are 
$g(\pm 1) = 0$. The variational formulation becomes: Find ${u} \in 
\N{V}_N$, ${g} \in \D{V}_N$ and ${f} \in \D{V}_N$ such that
\begin{align}
	\frac{\partial }{\partial t} \left< \nabla^2 u, \N{\psi}\right>_w &= 
	\left<h_u, \N{\psi} \right>_w + \nu \left<\nabla^4u, \N{\psi}\right>_w 
	&\forall \N{\psi} \in \N{V}_N, \label{eq:u1} \\
	\frac{\partial}{\partial t}\left<g, \D{\psi}\right>_w &= \left<h_g, 
	\D{\psi}\right>_w + \nu 
	\left<\nabla^2 g, \D{\psi}\right>_w &\forall \D{\psi} \in \D{V}_N ,
	\label{eq:g1} \\
	\left<f, \D{\psi}\right>_w &= \left<\frac{\partial u}{\partial x}, 
	\D{\psi}\right>_w &\forall \D{\psi} \in \D{V}_N. \label{eq:f1}
\end{align}
The two remaining velocity components are computed by projection to the 
Dirichlet space $\D{V}_N$: Find 
${v}, {w}$ in $\D{V}_N \times \D{V}_N$ such that
\begin{align}
\left<f, \D{\psi}\right>_w &= \left<\frac{\partial v}{\partial y} + 
\frac{\partial w}{\partial z}, \D{\psi}\right>_w \, &\forall \D{\psi} \in 
\D{V}_N, \label{eq:f2} \\
\left<g, \D{\psi}\right>_w &= \left<\frac{\partial w}{\partial y}  - 
\frac{\partial v}{\partial z}, \D{\psi}\right>_w \, &\forall \D{\psi} \in 
\D{V}_N, \label{eq:g2}
\end{align}
which simplifies considerably because all the derivatives are in periodic 
directions. Written in spectral space Eqs. (\ref{eq:f2}) and (\ref{eq:g2}) 
become simply
\begin{align}
\hat{f}_{\bm{k}} &= \imath \underline{m}\, \hat{v}_{\bm{k}} + \imath 
\underline{n}\, \hat{w}_{\bm{k}} &\forall \bm{k} \in \D{K}_N^0, \label{eq:f3} \\
\hat{g}_{\bm{k}} &= \imath \underline{m}\, \hat{w}_{\bm{k}} - \imath 
\underline{n}\, \hat{v}_{\bm{k}} & \forall \bm{k} \in \D{K}_N^0, \label{eq:g3}
\end{align}
where $\D{K}_N^0$ is used to denote that these equations can be solved for all 
wavenumbers except $m=n=0$. For $m=n=0$ we solve instead the 
corresponding momentum equations in $y$ and $z$ directions:
\begin{align}
\frac{\partial }{\partial t} \D{B}_{lq}\hat{v}(q, 0, 0, t) &= 
\D{B}_{lq}\hat{\mathcal{H}}_y(q, 0, 0, t) + \nu \D{A}_{lq} \hat{v}(q, 0, 0, t) 
& \forall\, l \in [0, 1, 
\ldots, N_x-2], \label{eq:v0}\\
\frac{\partial }{\partial t} \D{B}_{lq}\hat{w}(q, 0, 0, t) &= 
\D{B}_{lq}\hat{\mathcal{H}}_z(q, 0, 0, t) + \nu \D{A}_{lq} \hat{w}(q, 0, 0, t) 
& \forall\, l \in [0, 1, \ldots, N_x-2].\label{eq:w0}
\end{align}

Equations (\ref{eq:u1}-\ref{eq:f1}) are also solved in spectral space. 
Inserting for (\ref{eq:u_solx}) and a similar expansion for $g$ in $\D{V}_N$, 
the required inner products become
\begin{align}
\left<\nabla^4u, \N{\psi}_{\bm{k}}\right>_w &= h\left[ \left( 
\N{\phi}_q^{''''}, 
\N{\phi}_l\right)_w -2(\underline{m}^2+\underline{n}^2) \left( \N{\phi}_q^{''}, 
\N{\phi}_l\right)_w + (\underline{m}^2+\underline{n}^2)^2\left( \N{\phi}_q, 
\N{\phi}_l\right)_w  \right] \hat{u}_q, \\
\left< \nabla^2 u, \N{\psi}_{\bm{k}}\right>_w &= h\left[\left( \N{\phi}_q^{''}, 
\N{\phi}_l\right)_w - (\underline{m}^2+\underline{n}^2)\left( \N{\phi}_q, 
\N{\phi}_l \right)_w \right] \hat{u}_q, \\
\left< \nabla^2 g, \D{\psi}_{\bm{k}}\right>_w &= h\left[\left( \D{\phi}_q^{''}, 
\D{\phi}_l\right)_w - (\underline{m}^2+\underline{n}^2)\left( \D{\phi}_q, 
\D{\phi}_l \right)_w \right] \hat{g}_q, \\
\left<\frac{\partial u}{\partial x}, \D{\psi}_{\bm{k}}\right>_w &=
h\left(\N{\phi}_q^{'}, \D{\phi}_l\right)_w \hat{u}_q, \\
\left<h_u, \N{\psi} \right>_w &= h\left[-(\underline{m}^2+\underline{n}^2) 
\N{\mathcal{S}}(\mathcal{H}_x) - \imath \underline{m}\left(\D{\phi}_q^{'}, 
\N{\phi}_l \right)_w \hat{\mathcal{H}}_{y, q} - \imath 
\underline{n}\left(\D{\phi}_q^{'}, \N{\phi}_l \right)_w 
\hat{\mathcal{H}}_{z, q}\right], 
\\
\left< h_g, \D{\psi} \right>_w &= h\left[ \imath \underline{m}\, \D{B} 
\hat{\mathcal{H}}_z - \imath \underline{n}\, \D{B} \hat{\mathcal{H}}_{y} 
\right],
\end{align}
where for brevity in notation (as before) it is understood that the matrices 
act along the first dimension of the transformed variables, i.e.,  $ 
\left(\N{\phi}_q^{'}, \D{\phi}_l\right)_w \hat{u}_q$ is short for the matrix 
vector product $ \sum_{q=0}^{N_x-4}\left( \N{\phi}_q^{'}, \D{\phi}_l\right)_w 
\hat{u}(q, {m}, {n}, t)$. The inner products are used to set up linear systems 
of equations for the inhomogeneous wall normal direction. All inner products 
$\left(\cdot, \cdot\right)_w$ can be represented by sparse matrices.
%\begin{align}
%\frac{\partial }{\partial t} \left( \N{\psi}_j^{''}, \N{\psi}_k\right)_w 
%\hat{u}_j &=  \left(h_u, \N{\psi}_k \right)_w + \nu \left(\N{\psi}^{''''}_j, 
%\N{\psi}_k\right)_w \hat{u}_j, \, &\forall \N{\psi}_k \in \N{V}_N, \\
%\frac{\partial}{\partial t}\left(\D{\psi}_j, \D{\psi}_k\right)_w \hat{g}_j &= 
%\left(h_g, \D{\psi}_k\right)_w + \nu \left(\D{\psi}_j^{''}, 
%\D{\psi}_k\right)_w 
%\hat{g}_j, \, &\forall \D{\psi}_k \in \D{V}_N \\
%\left(\D{\psi}_j, \D{\psi}_k\right)_w \hat{f}_j &= \left(\N{\psi}_j^{'}, 
%\D{\psi}_k\right)_w \hat{u}_j, \, &\forall \D{\psi}_k \in \D{V}_N.
%\end{align}
The mass matrices, stiffness matrices $\N{A}_{lq} = \left( \N{\phi}_q^{''}, 
\N{\phi}_l\right)_w, 
\D{A}_{lq} = \left( \D{\phi}_q^{''}, \D{\phi}_l\right)_w 
$ and $\N{Q}_{lq} = \left(\N{\phi}^{''''}_q, \N{\phi}_l\right)_w$ are given by 
Shen, whereas $\left(\N{\phi}_q^{'}, \D{\phi}_l\right)_w$ and 
$\left(\D{\phi}_q^{'}, \N{\phi}_l\right)_w$ are given in the 
appendix.

It remains to discretize the equations in time. Kim, Moin and Mansour use 
Crank-Nicolson for the diffusive terms and Adams-Bashforth for the nonlinear 
convective terms, whereas a third order Runge-Kutta method is used for the 
nonlinear term by Moser, Kim and Mansour (1999). We have implemented both 
approaches, but for simplicity only the first one is described here. 
Discretization in time is performed using a constant time step $\triangle t$, 
such that the time is represented discretely as $t_n = n \triangle t, \, n \in 
\mathcal{Z}^{\ge}$. After 
discretizing in time, equations (\ref{eq:u1}) and (\ref{eq:g1}) are transformed 
to a matrix form that can be solved for $\hat{u}^{n+1}$ and $\hat{g}^{n+1}$ 
given wavenumbers $m$ and $n$
\begin{align}
\N{H}\hat{u}^{n+1} & = \left(2\N{A} - 2\underline{z}^2\N{B} - \N{H} 
\right)\hat{u}^{n} + \N{\mathcal{S}}(h_v^{n+1/2}) \triangle t & 
\forall \, {\bm{k}} \in \N{K}_N, 
\label{eq:ufin}\\ 
\D{H} \hat{g}^{n+1} &= \left(2 \D{B}-\D{H}\right) 
\hat{g}^{n} + \D{\mathcal{S}}(h_g^{n+1/2}) \triangle t &\forall \, 
{\bm{k}} \in 
\D{K}_N. \label{eq:gfin}
\end{align}
The convective terms are computed with 
a forward Adams-Bashforth projection, such that $\N{\mathcal{S}}(h_v^{n+1/2}) = 
\N{\mathcal{S}}(1.5h_v^{n} - 0.5 h_v^{n-1})$ and $\D{\mathcal{S}}(h_g^{n+1/2}) 
= \D{\mathcal{S}}(1.5h_g^{n} - 0.5 h_g^{n-1})$ and the coefficient matrices are 
given as
\begin{align}
\N{H}_{lq}(m, n) &= -\frac{\nu \triangle t}{2}\N{Q}_{lq} + \left( 1 + \nu 
\triangle t 
\underline{z}^2 \right) \N{A}_{lq} - \frac{2\underline{z}^2 + \nu \triangle t 
\underline{z}^4}{2} \N{B}_{lq} &\forall m, n \in K_N^p, \\
\D{H}_{lq}(m, n) &= -\frac{\nu \triangle 
t}{2}\D{A}_{lq} + (1 + \frac{\nu \triangle t \underline{z}^2}{2}) \D{B}_{lq} 
&\forall m, n \in K_N^p,
\end{align}
where $\underline{z}^2 = \underline{m}^2 + \underline{n}^2$. Equations 
(\ref{eq:v0}) and (\ref{eq:w0}) are also discretized in time using 
Crank-Nicolson and Adams-Bashforth, and we obtain
\begin{align}
\left(-\frac{\nu \triangle t}{2}\D{A} + \D{B} \right)\hat{v}_0^{n+1} &= 
\left(\frac{\nu \triangle t}{2}\D{A} + \D{B} 
\right)\hat{v}^{n}_0 + \triangle t\D{B}\hat{\mathcal{H}}_{y,0}^{n+1/2} , 
\label{eq:v00} \\
\left(-\frac{\nu \triangle t}{2}\D{A} + \D{B} \right)\hat{w}_0^{n+1} &= 
\left(\frac{\nu \triangle t}{2}\D{A} + \D{B} 
\right)\hat{w}^{n}_0 + \triangle t\D{B}\hat{\mathcal{H}}_{z,0}^{n+1/2}, 
\label{eq:w00}
\end{align}
where we have used notation $\hat{v}^{n}_0 = \hat{v}(l, 0, 0, t_n)$, 
$\hat{\mathcal{H}}^{n}_{y,0} = \hat{\mathcal{H}}_y(l, 0, 0, t_n)$ and similar 
for $\hat{w}^n_0$ and $\hat{\mathcal{H}}^n_{z,0}$.

\section{Implementation}

The solution procedure for the numerical method described in 
Sec.~\ref{sec:discretizationNS} is given in Algorithm \ref{alg:NS}

\begin{algorithm}
	\caption{Solution procedure for Navier Stokes equations}
    \label{alg:NS}
    \begin{algorithmic}[1]
    	\State \text{Initialize} $\bm{u}(\bm{x}, 0)$ and compute 
    	$\hat{g}_{\bm{k}}(0)$ and $\hat{u}_{\bm{k}}(0)$
		\State \text{Compute LU decomposition of} $\N{H}(m,n), \D{H}(m,n) 
		\,\forall \, m, n \in K_N^p$
		\State $t=0; n=0$
    	\While{$t<T$}
	    	\State \text{Compute nonlinear convection} $\bm{\mathcal{H}}^n$
	    	\State \text{Compute right hand sides of Eqs.~}(\ref{eq:ufin}), 
	    	(\ref{eq:gfin})
		    \State \text{Solve Eq.}(\ref{eq:ufin}) for 
		    $\hat{u}^{n+1}_{\bm{k}} \, \forall \,\bm{k} \in \N{K}_N$
		    \State \text{Solve Eqs.}(\ref{eq:gfin}) for 
		    $\hat{g}^{n+1}_{\bm{k}} \, \forall\, \bm{k} \in \D{K}_N$
		    \State \text{Solve Eqs.}(\ref{eq:f3}, \ref{eq:g3}) for  
		    $\hat{v}_{\bm{k}}, \hat{w}_{\bm{k}}\, \forall\, \bm{k} 
		    \in \D{K}_N^0$
		    \State \text{Solve Eqs.}(\ref{eq:v00}, \ref{eq:w00}) for 
		    $\hat{v}_0^{n+1}, \hat{w}_0^{n+1}\, 
		    \forall \, l \in [0, 1, \ldots, N_x-2] $
		    \State $n \gets n + 1$
		    \State $t \gets t+\triangle t$
		    \State \text{Update to new time step}
	    \EndWhile

    \end{algorithmic}
\end{algorithm}

The major computational cost comes from  computing the nonlinear convection 
and solving for Eqs.~(\ref{eq:ufin}, \ref{eq:gfin}). The nonlinear 
convection 
term $\bm{\hat{\mathcal{H}}}$ is computed using a pseudo-spectral method with 
the 3/2-rule for de-aliasing in the periodic directions (ref Canuto). The 
transforms are carried out using fast Fourier and fast Chebyshev transforms 
that are all of order $O(N\log N)$ per one-dimensional transform of size 
N. The Shen transforms make use of the Chebyshev transforms that are using 
discrete cosine transforms. The forward scalar product is implemented as shown 
in Algorithm \ref{alg:fst}. A complete transform is obtained through a scalar 
product followed by solving the linear system involving the mass 
matrix, just like shown in Eq. (\ref{eq:fstB}) for space $\N{V}_N$ and similar 
for $\D{V}_N$ and $V_N$. Since the mass matrices are diagonal, tridiagonal and 
pentadiagonal for $V_N, \D{V}_N$ and $\N{V}_N$ respectively, solving the linear 
system is $O(N)$ and the complete transform still scales like $O(N\log N)$. 
\begin{algorithm}
	\caption{Forward scalar product for all spaces $V_N, \D{V}_N, \N{V}_N$. 
	Here "dct" is the discrete cosine transform from SciPy.}
	\label{alg:fst}
	\begin{algorithmic}[1]
	\Procedure{fastScalar}{}
		\State \textbf{Input:} $\{f_j\}_{j=0}^{N}$
		\State \textbf{Input:} points
		\State \textbf{Input:} space
		\If{ points = "Gauss-Lobatto"}
			\State $\{ w_k\}_{k=0}^{N} \gets \text{dct}(f, \text{type=2}, 
			\text{axis=0})\frac{\pi}{2 N}$
		\Else{ points = "Gauss-Chebyshev"}
		    \State $\{ w_k\}_{k=0}^{N} \gets \text{dct}(f, \text{type=1}, 
		    \text{axis=0}) \frac{\pi}{2 (N-1)}$
	    \EndIf
	    \If{ space = $\D{V}_N$}
	        \For{$k=0$ to $N-2$}
		        \State $\hat{f}_k \gets w_k - w_{k+2}$
	        \EndFor
			\State \Return $\{\hat{f}_k\}_{k=0}^{N-2}$	        
	    \ElsIf{ space = $\N{V}_N$}	                
  	        \For{$k=0$ to $N-4$}
       	        \State $\hat{f}_k \gets w_k - \frac{2(k+2)}{k+3} w_{k+2} + 
       	        \frac{k+1}{k+3}w_{k+4}$
   	        \EndFor
   	        \State \Return $\{\hat{f}_k\}_{k=0}^{N-4}$
   	    \Else{}
   	        \For{$k = 0$ to $N$}
     	        \State $f_k \gets \frac{2}{\pi} w_k $
     	    \EndFor
     	    \State $f_0 \gets f_0 / 2$
     	    \If{ points = "Gauss-Lobatto"}
	     	\State $f_N \gets f_N / 2$	     	
     	    \EndIf
			\State \Return $\{f_k\}_{k=0}^{N}$    
	    \EndIf
	\EndProcedure
\end{algorithmic}
\end{algorithm}

We will now describe very efficient direct solvers for Eqs.~(\ref{eq:ufin}, 
\ref{eq:gfin}) that are of order $O(N)$. The solvers are generated using clever 
direct LU decomposition and only a few diagonals need to be stored for each 
final lower (L) and upper (U) triangular matrix.

To solve Eq. (\ref{eq:gfin}) we perform an LU decomposition of the coefficient 
matrix $\D{H}$. $\D{H}$ consists of one subdiagonal, and every 
second upper diagonal is zero. As such, we may split the matrix into two 
smaller (and decoupled) matrices for odd and even coefficients $\D{H}^e_{k,j} = 
\D{H}_{2k,2j}$ and $\D{H}^o_{k,j} = \D{H}_{2k+1,2j+1}$, where the two matrices 
$\D{H}^{e,o}$ are of type upper 
Hessenberg, with only 4 distinct values on each row. This structure of $\D{H}$ 
allows us to perform a very efficient tailored LU decomposition, as shown in 
Algorithm (\ref{alg:lu}), and a very efficient tailored solve through 
Algorithm (\ref{alg:lusolve}). The efficiency follows from the fact that 
$U^{o,e}_{kj} = U^{o,e}_{kk+2} \, \forall \, j > k+2$, i.e., the upper 
triangular matrix has a maximum of three distinct values for each row, which 
allows for a very fast solve through back substitution. 
Regarding Algorithm 
(\ref{alg:lusolve}), we note that the ability to decouple the system into odd 
and even coefficients leads to two subsystem that may be trivially solved 
simultaneously in two threads by letting one thread call LUSolveoddeven with 
even coefficients, and the second thread call it for odd. For optimal 
performance, the odd and even coefficient would then need to be stored 
contiguously in computer memory, and not alternately, which would be the normal 
way of storing the coefficients.

\begin{algorithm}
\caption{LU factorization of matrix $H_{k,j}$, that may be split into two 
decoupled upper Hessenberg matrices $H^e_{k,j} = H_{2k,2j}$ and $H^o_{k,j} = 
H_{2k+1,2j+1}$ for even and odd coefficients. Return LU factorization for even 
and odd coefficients, such that $L^eU^e=H^e$ and $L^oU^o=H^o$. Note that 
$L^{e}$ and $L^{o}$ each have one single nonzero diagonal in addition to the 
unity main diagonal, and as such only a vector is stored. Each $U^{e}$ and 
$U^{o}$ has three distinct diagonals and $U^{e}_{kj}=U^{e}_{k, k+2} \,\forall 
j 
> k+2, j< M^e$ and $U^{o}_{kj}=U^{o}_{k, k+2} \,\forall j > k+2, j< M^o$. As 
such, a sparse three-diagonal storage may be used for $U^e$ and $U^o$.}
\label{alg:lu}
\begin{algorithmic}[1]
\Procedure{LUoddeven}{}
  \State \textbf{Input:} $\{H\}_{k,j=0}^{N-2}$
  \State \textit{integer} $M^e \gets (N-2)/2$
  \State \textit{integer} $M^o \gets (N-3)/2$
  \State ${U}^e_{0j} \gets {H}_{0,2j}, \, \forall j=0,1,2 $
  \State ${U}^o_{0j} \gets {H}_{1,2j+1}, \, \forall j=0,1,2 $
  \For{$k=1, 2,  \ldots, M^e-2$}
    \State ${L}^e_{k} \gets \frac{{H}_{2k,2(k-1)}}{{U}^e_{k-1,k-1}}$
    \State ${U}^e_{kj} \gets {H}_{2k,2j} - {L}^e_{k}{U}^e_{k-1,j}, \, \forall 
j=k, k+1, k+2$
    \State ${L}^o_{k} \gets \frac{{H}_{2k+1,2k-1}}{{U}^o_{k-1,k-1}}$
    \State ${U}^o_{kj} \gets {H}_{2k+1,2j+1} - {L}^o_{k}{U}^o_{k-1,j}, \, 
\forall j=k, k+1, k+2$
  \EndFor

  \State $k \gets M^e-1$
  \State ${L}^e_{k} \gets \frac{{H}_{2k,2(k-1)}}{{U}^e_{k-1,k-1}}$
  \State ${U}^e_{kj} \gets {H}_{2k,2j} - {L}^e_{k}{U}^e_{k-1,j}, \, \forall 
j=k, k+1$
  \State ${L}^o_{k} \gets \frac{{H}_{2k+1,2k-1}}{{U}^o_{k-1,k-1}}$
  \State ${U}^o_{kj} \gets {H}_{2k+1,2j+1} - {L}^o_{k}{U}^o_{k-1,j}, \, 
\forall j=k, k+1$
  \State $k \gets M^e$
  \State ${L}^e_{k} \gets \frac{{H}_{2k,2(k-1)}}{{U}^e_{k-1,k-1}}$
  \State ${U}^e_{kk} \gets {H}_{2k,2k} - {L}^e_{k}{U}^e_{k-1,k}$
  \If{$M^o == M^e$ }
    \State ${L}^o_{k} \gets \frac{{H}_{2k+1,2k-1}}{{U}^o_{k-1,k-1}}$
    \State ${U}^o_{kk} \gets {H}_{2k+1,2k+1} - {L}^o_{k}{U}^o_{k-1,k}$

  \EndIf  
  \State \Return $\{L^e_k\}_{k=0}^{M^e},\{U^e\}_{k,j=0}^{M^e}, 
\{L^o_k\}_{k=0}^{M^o},\{U^o\}_{k,j=0}^{M^o}$
\EndProcedure
\end{algorithmic}
\end{algorithm}

\begin{algorithm}
\caption{Solve of $H_{kj}u_j = b_k$, where $H$ has previously been LU 
factorized for odd even coefficients using Algorithm (\ref{alg:lu}).}
\label{alg:lusolve}
\begin{algorithmic}[1]
\Procedure{LUSolveoddeven}{}
  \State \textbf{Input:} $\{U_{kj}\}_{k,j=0}^{M}, \{L_k\}_{k=0}^{M}, 
\{b_k\}_{k=0}^{M}$ 
  \State // Solve $Ly=b$ by forward elimination 
  \State $y_0 \gets b_0$
  \For{$k = 1$ \textbf{to} $M$}
    \State $y_k \gets b_k - L_ky_{k-1}$
  \EndFor
  \State // Solve $Uu=y$ with back substitution
  \State double $s = 0.0$     
  \State $u_M = y_M / U_{MM}$    
  \For{$i = M-1$ \textbf{to} $0$ \textbf{step} $-1$}
     \State $u_i \gets y_i - U_{i, i+1}\cdot u_{i+1}$
     \If{$i < M-1$}
       \State $s \gets s + u_{i+2}$
       \State $u_i \gets u_i - s\cdot U_{i,i+2}$            
     \EndIf
     \State $u_i \gets  u_i / U_{i,i}$
   \EndFor
  \State \Return $\{u_k\}_{k=0}^{M}$  
\EndProcedure
\end{algorithmic}

\begin{algorithmic}[1]
\Procedure{LUSolve}{}
  \State \textbf{Input: $\{u_k\}_{k=0}^{N-2}, \{b_k\}_{k=0}^{N-2}, 
\{U_{kj}^{e}\}_{k,j=0}^{M^{e}}, \{U_{kj}^{o}\}_{k,j=0}^{M^{o}}, 
\{L^{e}_k\}_{k=0}^{M^{e}}, \{L^{o}_k\}_{k=0}^{M^{o}}$}
  \State $\{u_{2k}\}_{k=0}^{M^e} \gets $ LUSolveoddeven($U^{e}, L^e, 
\{b_{2k}\}_{k=0}^{M^e}$)
  \State $\{u_{2k+1}\}_{k=0}^{M^o} \gets $ LUSolveoddeven($U^{o}, L^o, 
\{b_{2k+1}\}_{k=0}^{M^o}$)
  \State \Return $\{u_k\}_{k=0}^{N-2}$
\EndProcedure
\end{algorithmic}
\end{algorithm}

The biharmonic operator $\N{H}$ is more challenging to solve efficiently, but 
it is still possible to find an algorithm that is $O(N)$ for a system of $N$ 
unknowns. Note 
that $\N{H}$ is the sum of three matrices $\N{Q}, \N{A}$ and $\N{B}$ and can be 
written as
\begin{equation}
\N{H} = c_0\N{Q} + c_1\N{A} + c_2 \N{B},
\end{equation}
where $c_0, c_1$ and $c_2$ are constants. The matrix $\N{A}_{kj}$ contains only 
three nonzero diagonals at $ j = k-2, k, k+2$,  whereas matrix $\N{B}_{kj}$ 
contains five 
nonzero diagonals at $ j = k-4, k-2, k, k+2, k+4$. The nonzero elements of the 
$\N{Q}_{kj}$ matrix are given as
\begin{align}
 \N{Q}_{kk} &= -4(k+1)(k+2)^2 \\
 \N{Q}_{kj} &= p_kq_j + r_ks_j, \quad j = k+2, k+4, \ldots N_x-4
\end{align}
where 
\begin{align}
p_k &= 8 k (k+1)(k+2)(k+4)\pi, &q_j = \frac{1}{j+3} \label{eq:pk} \\
r_k &= 24(k+1)(k+2)\pi, &s_j =  \frac{(j+2)^2}{(j+3)} \label{eq:rk}
\end{align} 
For each row in $\N{H}_{kj}$ we have that $\N{H}_{kj}=c_0\N{Q}_{kj}\, \forall j 
> k+4$. A direct LU decomposition of $\N{H}$ can be performed as shown in Alg. 
\ref{alg:lu_biharmonic}, which leads to a lower triangular 
matrix $\N{L}_{kj}$ with two nonzero diagonals at $j=k-2$ and $k-4$ plus a 
unity 
main diagonal. The upper triangular matrix $\N{U}_{kj}$ contains three distinct 
diagonals at $j=k, k+2$ and $k+4$. The remaining part of $\N{U}_{kj}$ can be 
written as
\begin{equation}
\N{U}_{kj} = a_k q_j + b_k s_j, \quad j = k+6, k+8, \ldots, N_x-4 \label{eq:ab}
\end{equation}
where $\bm{a}$ and $\bm{b}$ are two new vectors that can be computed 
recursively as shown in Alg. (\ref{alg:ab}). If $\bm{a}$ and $\bm{b}$ are 
pre-computed 
the total storage requirement for the complete LU decomposition is no more than 
$7N$ and the solution of the complete system can be obtained very quickly with 
one forward elimination and a back solve. The back 
solve is very fast because of the recursive formulation (\ref{eq:ab}), leading 
to a formula like (valid for $k \le N_x-8$)
\begin{equation}
u_k = \left(y_k - \N{U}_{k, k+2} u_{k+2} - \N{U}_{k, k+4} u_{k+4} - 
a_k\sum_{\substack{j=k+6\\k-j \text{ even}}}^{N_x-4} q_j u_j - 
b_k\sum_{\substack{j=k+6\\k-j \text{ even}}}^{N_x-4} 
s_j u_j\right)/\N{U}_{kk},
\end{equation}
that only requires 
one new addition to the row-sums for each back-traversed
row, see Alg. (\ref{alg:SolveBiharmonic}).

\begin{algorithm}
	\caption{LU decomposition of biharmonic operator $\N{H}$. Current algorithm 
	is using dense storage of $U_{kj}$, but no more than the 3 diagonals need 
	to be stored.}
	\label{alg:lu_biharmonic}
	\begin{algorithmic}[1]
		\Procedure{LUbiharmonic}{}
			\State \textbf{Input:} $\{\N{H}_{kj}\}_{k, j = 0}^{N}$
	        \State $U_{0j} = \N{H}_{0j} \quad \forall \, j \in 0, 2, \ldots, N$
	        \State $U_{1j} = \N{H}_{1j} \quad \forall \, j \in 1, 3, \ldots, N$
	        \State $L_{20} = \N{H}_{20}/U_{00}$
	        \State $L_{31} = \N{H}_{31}/U_{11}$
	        \State $U_{2j} = L_{20} U_{0j} \quad \forall \, j \in 2, 4, \ldots, 
	        N$
	        \State $U_{3j} = L_{31} U_{1j} \quad \forall \, j \in 3, 5, \ldots, 
	        N$
	        \For{$k = 4, 5, \ldots, N$}
	            \State $L_{k,k-2} = {U}_{k,k-2}/{U}_{k-2, k-2}$
		        \State $U_{kj} = {U}_{kj} - L_{k, k-2} U_{k-2, j} \quad 
		        \forall j = k, k+2, \ldots, N$
		        \If{$k <= N-2$}
		        \State $L_{k+2,k-2} = \N{H}_{k+2, k-2} / U_{k-2, k-2}$
		        \State $U_{k+2,j} = \N{H}_{k+2,j} - L_{k+2, k-2} U_{k-2, j} 
		        \quad \forall j = k-2, k, \ldots, N$
		        \EndIf
	        \EndFor
	        \State \Return $\{L_{kj}\}_{k, j = 0}^{N}$, $\{U_{kj}\}_{k, j = 
	        0}^{N}$
		\EndProcedure	
	\end{algorithmic}
\end{algorithm}

\begin{algorithm}
	\caption{Recursive formula to compute $\{a_k\}_{k=0}^{N_x-4}, 
	\{b_k\}_{k=0}^{N_x-4}$ in Eq. (\ref{eq:ab}). Note that the lower triangular 
	matrix $\{L_{kj}\}_{k,j=0}^{N_x-4}$ contains only two nonzero diagonals and 
	the parameters $p_k$ and $r_k$ are given in Eqs. (\ref{eq:pk}) and 
	(\ref{eq:rk}) respectively. }
    \label{alg:ab}
	\begin{algorithmic}[1]
		\Procedure{ReduceAB}{}
		\State \textbf{Input:} $ \{L_{kj}\}_{k,j=0}^{N_x-4}$
		\For{$k=0,1$}
		     \State $a_k = p_k$
		     \State $b_k = r_k$
		\EndFor
		\For{$k=2,3$}
		    \State $a_k = p_k - L_{k,k-2}\,a_{k-2}$
		    \State $b_k = r_k - L_{k, k-2}\,b_{k-2}$
		\EndFor
		\For{$k=4, 5, \ldots, N_x-4$}
     		\State $a_k = p_k - L_{k,k-2}\,a_{k-2} - L_{k, k-4} \,a_{k-4}$
	    	\State $b_k = r_k - L_{k, k-2}\,b_{k-2}- L_{k, k-4} \,b_{k-4}$
		\EndFor
		\State \Return $\{a_k\}_{k=0}^{N_x-4}, \{b_k\}_{k=0}^{N_x-4}$
        \EndProcedure
\end{algorithmic}
\end{algorithm}

\begin{algorithm}
	\caption{Solve biharmonic Eq. (\ref{eq:ufin}) with pre-computed $L$, $U$ 
	matrices. The parameters $q_j$ and $s_j$ are given in Eqs. (\ref{eq:pk}) 
	and 
	(\ref{eq:rk}) respectively. }
	\label{alg:SolveBiharmonic}
	\begin{algorithmic}[1]
		\Procedure{LUSolveBiharmonic}{}
		  \State \textbf{Input:} $\{U_{kj}\}_{k,j=0}^{N}, 
		  \{L_{kj}\}_{k,j=0}^{N}, 
		  \{f_k\}_{k=0}^{N}, \{a_k\}_{k=0}^{N}, \{b_k\}_{k=0}^{N}$ 
		  \State // Solve $Ly=f$ by forward elimination 
		  \State $y_0 \gets f_0$
		  \State $y_1 \gets f_1$
		  \State $y_2 \gets f_2 - L_{20}y_{0}$
		  \State $y_3 \gets f_3 - L_{31}y_{1}$
		  \For{$k = 4$ \textbf{to} $N$}
		  \State $y_k \gets f_k - L_{k,k-2}y_{k-2} - L_{k, k-4}y_{k-4}$
		  \EndFor
		  \State // Solve $Uu=y$ with back substitution                    
		  \State $u_N = y_N / U_{NN}$
	      \State $u_{N-1} = y_{N-1} / U_{N-1,N-1}$
	      \State $u_{N-2} = \left(y_{N-2} - U_{N-2, N}y_N\right)/U_{N-2,N-2}$
	      \State $u_{N-3} = \left(y_{N-3} - U_{N-3, 
	      N-1}y_{N-1}\right)/U_{N-3,N-3}$

	      \State $q^o = q^e = s^o = s^e = 0.0$
		  \For{$k = N-4$ \textbf{to} $0$ \textbf{step} $-1$}
		  \State $u_k \gets y_k - U_{k, k+2} u_{k+2} - U_{k, k+4} 
		  u_{k+4} $ 
		  \If{$k < N-5$}
		  \State $j=k+6$
		  \If{$k$ is odd}
		    \State $q^o \gets q^o + u_{j}q_{j}$
		    \State $s^o \gets s^o + u_{j}s_{j}$
		    \State $u_k \gets u_k - c_0 q^o a_k - c_0 s^o b_k $	    
		  \Else
		    \State $q^e \gets q^e + u_{j}q_{j}$
		    \State $s^e \gets s^e + u_{j}s_{j}$	  
		    \State $u_k \gets u_k - c_0 q^e a_k - c_0 s^e b_k $
		  \EndIf
		  \EndIf
		  \State $u_k \gets  u_k / U_{kk}$
		  \EndFor
		  \State \Return $\{u_k\}_{k=0}^{N}$

		\EndProcedure
	\end{algorithmic}
\end{algorithm}

%\section{The incremental pressure correction scheme}
%
%The Navier Stokes equations are discretized in time using a second order 
%incremental pressure correction scheme with Crank-Nicolson on diffusion and 
%explicit Adams Bashforth on the nonlinear convection. The incremental pressure 
%correction scheme (IPCS) that is used consists of a tentative momentum 
%equation 
%followed by a pressure correction. The two steps may be performed iteratively. 
%The equations are iterated forward in time in equal size time steps $\triangle 
%t$ from initial conditions $u(\bm{x}, 0) = u^0(\bm{x})$ at $t=0$, such that 
%$t_n = n\triangle t$, $n \in \mathcal{Z}$. For turbulent channel flows initial 
%conditions are simply a perturbed state used to get a turbulent flow going. 
%
%The IPCS attempts to find velocity vector $\bm{u}^{n}$ on time step $t_n$  and 
%pressure midway between $t_n$ and $t_{n-1}$, $p^{n-1/2}$, given the solutions 
%on all previous time steps. The three steps that are required solved on each 
%time step are
%\begin{align}
%i)&\,\, \frac{\bm{u}^{*}-\bm{u}^{n-1}}{\triangle t} + \bm{N}^{n-1/2}   = \nu 
%\nabla^2 \bm{u}^{n-1/2} - \nabla{p}^{n-3/2} + \bm{f}, \quad \bm{u}^*=0 \text{ 
%for } x=\pm1   \label{eq:NSCN} \\
%ii)&\,\,\begin{cases}
%  \frac{\bm{u}^{n}-\bm{u}^*}{\triangle t} &= -\nabla p^* \\
%  \nabla \cdot \bm{u}^{n} &= 0, \quad \bm{u}^n\cdot \bm{n} = 0,\text{ for } 
%x=\pm1
%  \end{cases} \label{eq:momupdate}\\
%  iii)&\,\,p^{n-1/2}=p^{n-3/2}+p^* \label{eq:pupdate} 
%\end{align}
%Here $\bm{u}^{n-1/2} = 0.5(\bm{u}^*+\bm{u}^{n-1})$, $\bm{u}^*$ is a tentative 
%velocity, $p^*$ is a pressure correction, and the explicit convection is 
%computed as $\bm{N}^{n-1/2}=1.5(\bm{u}^{n-1}\cdot \nabla) \bm{u}^{n-1} - 
%0.5(\bm{u}^{n-2}\cdot \nabla) \bm{u}^{n-2}$. The second step is computed by 
%solving a Poisson equation for the pressure correction
%\begin{equation}
%  \nabla^2p^* = \frac{\nabla \cdot \bm{u}^*}{\triangle t}
%\end{equation}
%and subsequently updating the velocity through (\ref{eq:momupdate}).
%
%All three velocity components use the Dirichlet basis (\ref{eq:u_sol}), 
%whereas the pressure is searched for in $\N{V}_N$
%\begin{equation}
%p(\bm{x}, t) = \frac{1}{N_yN_z}\sum_{\bm{k} \in \N{K}_N} \hat{p}_{\bm{k}}(t) 
%\N{\psi}_{\bm{k}}(\bm{x}), \label{eq:p_sol}
%\end{equation}
% 
%The spectral Galerkin method becomes: 
%
%1) Find $\bm{u}^*$ in $\D{V}_N^3$ by solving
%\begin{multline}
% \frac{1}{\triangle t}\left<\bm{u}^*-\bm{u}^{n-1}, \bm{\D{\psi}}\right>_w + 
%\left<\bm{N}^{n-1/2}, \bm{\D{\psi}} \right>_w   = \nu \left<\nabla^2 
%\bm{u}^{n-1/2}, \bm{\D{\psi}} \right>_w \\ - \left<\nabla{p}^{n-3/2}, 
%\bm{\D{\psi}} \right>_w + \left<\bm{f}, \bm{\D{\psi}} \right>_w \,\,  \forall 
%\, \bm{\D{\psi}} \in \D{V}_N^3. \label{eq:tentative}
%\end{multline}
%
%2) Compute the pressure correction $p^*\in \N{V}_N$ by solving
%\begin{equation}
%\left< \nabla^2 p^*, \N{\psi} \right>_w = \frac{1}{\triangle t} \left<\nabla 
%\cdot \bm{u}^*, \N{\psi} \right>_w\,\, \forall \, \N{\psi} \in \N{V}_N. 
%\label{eq:pcorr}
%\end{equation}
%
%3) Update the velocity, $\bm{u}^n \in \D{V}_N^3$, and pressure, $p^{n-1/2} \in 
%\N{V}_N$, by solving
%\begin{align}
%\left< \bm{u}^n, \bm{\D{\psi}}\right>_w &= \left< \bm{u}^*, 
%\bm{\D{\psi}}\right>_w - \triangle t \left< \nabla p^*, \bm{\D{\psi}} 
%\right>_w 
%\,\,  &\forall \, \bm{\D{\psi}} \in \D{V}_N^3, \label{eq:u_update} \\
%\left<p^{n-1/2}, \N{\psi}\right>_w &= \left<p^{n-3/2}, \N{\psi}\right>_w + 
%\left<p^*, \N{\psi}\right>_w \,\, &\forall \, \N{\psi} \in 
%\N{V}_N,\label{eq:pres_update}
%\end{align}
%where the pressure update simply resorts to $\hat{p}^{n-1/2} = \hat{p}^{n-3/2} 
%+ \hat{p}^* \, \forall \, \bm{k}\in \N{K}_N$.
%
%\section{Implementation}
%The vector valued equations (\ref{eq:tentative}) and (\ref{eq:u_update}) are 
%solved in a segregated manner, one component at the time. We use $u_i$ and 
%$\hat{u}_i$ for component $i$ of the real and transformed velocity vector. The 
%inner products in Eqs. (\ref{eq:tentative}, \ref{eq:pcorr}, \ref{eq:u_update}, 
%\ref{eq:pres_update}) lead to a range of new matrices
%\begin{align}
% \left< u_i^* - u_i^{n-1}, \D{\psi}\right>_w &= 
%h\D{B}\left({\hat{u}}_i^*-{\hat{u}}_i^{n-1} \right) \quad &i \in (0,1,2), \\ 
% \left<\nabla^2 {u}_i^{n-1/2}, \D{\psi}\right>_w &= -h  \left( \D{A} 
%+(\underline{m}^2+\underline{n}^2)\D{B}\right) \hat{u}_i^{n-1/2} \quad &i \in 
%(0,1,2), \\ 
% \left< f_i, \D{\psi}\right>_w &= h\D{\mathcal{S}}(f_i) &i \in (0,1,2), \\
% \left<N_i^{n-1/2}, \D{\psi} \right>_w &= h\D{\mathcal{S}}(N_i^{n-1/2}) &i \in 
%(0,1,2), \\  
% \left< \nabla p^{n-3/2}, \bm{\D{\psi}} \right>_w &= h\left[\grave{C},\, 
%\imath \underline{m} \grave{B},\, \imath \underline{n} \grave{B} \right] 
%\hat{p}^{n-3/2}, \\ 
% \left< \nabla^2 p^*, \N{\psi} \right>_w &= -h\left(\N{A} +(\underline{m}^2 + 
%\underline{n}^2)\N{B} \right) \hat{p}^*, \\ 
% \left<\nabla \cdot \bm{u}^*, \N{\psi} \right>_w &= h\left( \tilde{C} 
%\hat{u}^* + \imath \underline{m} \tilde{B} \hat{v}^* + \imath \underline{n} 
%\tilde{B} \hat{w}^* \right), 
%\end{align}
%where $\D{A}_{kj} = -\left(\D{\phi}^{''}_{j}, \D{\phi}_{k} \right)_w, 
%\N{A}_{kj} = -\left(\N{\phi}^{''}_{j}, \N{\phi}_{k} \right)_w,  \grave{C}_{kj} 
%= \left(\N{\phi}^{'}_j, \D{\phi}_k \right)_w, \tilde{C}_{kj} = 
%(\D{\phi}^{'}_j, 
%\N{\phi}_k)_w, \N{B}_{kj} = \left( \N{\phi}_j, \N{\phi}_k \right)_w, 
%\grave{B}_{kj} = (\N{\phi}_j, \D{\phi}_k)_w$ and $\tilde{B}_{kj} = 
%(\D{\phi}_j, 
%\N{\phi}_k)_w$ are all more or less sparse matrices. Arranging the governing 
%equations on matrix form we obtain for the three tentative velocity components 
%in (\ref{eq:tentative})
%\begin{align}
%\D{H} \hat{u}^* &= \left(\frac{4}{\nu \triangle t} \D{B}-\D{H}\right) 
%\hat{u}^{n-1} - \left[\D{\mathcal{S}}(N_x^{n-1/2}) + \grave{C} \hat{p}^{n-3/2} 
%+ \D{\mathcal{S}}(f_x)\right]\frac{2}{\nu}, \label{eq:tentativeUA}\\
%\D{H} \hat{v}^* &=  \left(\frac{4}{\nu \triangle t} 
%\D{B}-\D{H}\right)\hat{v}^{n-1} - \left[\D{\mathcal{S}}(N_y^{n-1/2}) + \imath 
%\underline{m} \grave{B} \hat{p}^{n-3/2} + 
%\D{\mathcal{S}}(f_y)\right]\frac{2}{\nu}, \label{eq:tentativeVA}\\
%\D{H} \hat{w}^* &= \left(\frac{4}{\nu \triangle t} \D{B}-\D{H}\right) 
%\hat{w}^{n-1} - \left[\D{\mathcal{S}}(N_z^{n-1/2}) + \imath \underline{n} 
%\grave{B} \hat{p}^{n-3/2} + 
%\D{\mathcal{S}}(f_z)\right]\frac{2}{\nu},\label{eq:tentativeWA}
%\end{align}
%where $\D{H}=\D{A} +(\underline{m}^2 + \underline{n}^2 + \frac{2}{\nu 
%\triangle t})\D{B}$ is a matrix consisting of one subdiagonal, and where every 
%second diagonal is zero. As such, we may split the matrix into two smaller 
%(and 
%decoupled) matrices for odd and even coefficients $H^e_{k,j} = H_{2k,2j}$ and 
%$H^o_{k,j} = H_{2k+1,2j+1}$, where the two matrices $H^{e,o}$ are of type 
%upper 
%Hessenberg, with only 4 distinct values on each row. This structure of $H$ 
%allows us to perform a very efficient tailored LU decomposition, as shown in 
%Algorithm (\ref{alg:lu}), and a very efficient tailored solve through 
%Algorithm 
%(\ref{alg:lusolve}). Regarding Algorithm (\ref{alg:lusolve}), we note that the 
%ability to decouple the system into odd and even coefficients leads to two 
%subsystem that may be trivially solved simultaneously in two threads by 
%letting 
%one thread call LUODDEVEN with even coefficients, and the second thread call 
%it 
%for odd. 
%
%The linear system arising for the pressure correction equation is similar to 
%the tentative velocities
%\begin{equation}
%\N{H}\hat{p}^* = -\frac{1}{\triangle t}\left( \tilde{C} \hat{u}^* + \imath 
%\underline{m} \tilde{B} \hat{v}^* + \imath \underline{n} \tilde{B} \hat{w}^* 
%\right), \label{eq:pressureA}
%\end{equation}
%where $\N{H}=\N{A} +(\underline{m}^2 + \underline{n}^2)\N{B}$ is of the same 
%structure as $\D{H}$ and as such  should be factorizable in the same way. 
%However, entries of the matrix $\N{H}_{kj}$ contain the column index $j^2$ 
%(see 
%$\N{A}$), and as such the values on each row of the matrix $\N{H}_{kj}$ are 
%not 
%constant for $j \ge k+4$ and the LU decomposition of $\N{H}$ will require a 
%full upper $U$ matrix. To circumvent this we define $\N{H}_{kj} = 
%\underline{H}_{kl}J_{lj}$, where $J_{kj}= j^2\delta_{kj}$ (no summation 
%implied). We then set $w_k=J_{kj}\hat{p}_j$, use $b_k$ for the right hand side 
%of (\ref{eq:pressureA}) and solve
%\begin{align}
%  \underline{H}_{kj}w_j &= b_k &\forall \, k \in 1, 2, \ldots, N_x-2, \\
%  \hat{p}_k &= J_{kj}^{-1}w_j &\forall \, k \in 1, 2, \ldots, N_x-2.
%\end{align}
%Here the first system $\underline{H}_{kj} w_j = b_k$ can be solved using the 
%same LU factorization as the $\D{H}$ matrix, since 
%$\underline{H}_{kj}=\underline{H}_{k, k+4}\, \forall \, j=k+6, k+8, \ldots$ 
%and 
%there are only 4 distinct values on each row. The implementation is the same 
%as 
%for $\D{H}$, except that for the matrix $\underline{H}$ the indices start at 1 
%instead of zero.
%
%The velocity update is computed on matrix form as
%\begin{align}
%  \hat{u}^n &= \hat{u}^* - \triangle t \D{B}^{-1} \left( \grave{C} \hat{p}^* 
%\right),\\
%  \hat{v}^n &= \hat{v}^* - \triangle t \D{B}^{-1} \left( \imath \underline{m} 
%\grave{B} \hat{p}^* \right),\\ 
%  \hat{w}^n &= \hat{w}^* - \triangle t \D{B}^{-1} \left( \imath \underline{n} 
%\grave{B} \hat{p}^* \right).
%\end{align}
%using a tridiagonal solver.
%
%The convection term is computed in standard form with a pseudo-spectral method 
%that requires a fast scalar product
%\begin{equation}
%  \D{\mathcal{S}}(N_i^{n-1}) = \D{\mathcal{S}}(\bm{u}^{n-1} \cdot \nabla 
%u_i^{n-1}) \quad \forall \, i \in (0, 1, 2).
%\end{equation}
%We obtain the gradients of $u_i$ in real space by projecting $\partial 
%u/\partial x$ to $\D{V}_{N_x}$ (boundary condition $\partial u/\partial x(\pm 
%1)=0$ follows from continuity) and $\partial v/\partial x$ and $\partial 
%w/\partial x$ to $V_{N_x}$. 
%\begin{align}
%  \frac{\partial u}{\partial x} &= \D{\mathcal{T}}^{-1} 
%(\D{B}^{-1}(\D{C}\hat{u})), \\
%  \frac{\partial v}{\partial x} &= \mathcal{T}^{-1} 
%(B^{-1}(\acute{C}\hat{v})), \\
%  \frac{\partial w}{\partial x} &= \mathcal{T}^{-1} 
%(B^{-1}(\acute{C}\hat{w})),  
%\end{align}
%where $\D{C}_{kj} = (\D{\phi}_j^{'}, \D{\phi}_k)_w$,  $\acute{C_{kj}} = 
%(\D{\phi}_j^{'}, T_k)_w$ and $B=(T_j, T_k)_w$. Partial derivatives with 
%respect 
%to periodic directions are computed as
%\begin{align}
% \frac{\partial u_i}{\partial y} &= \D{\mathcal{T}}^{-1}(\imath \underline{m} 
%\hat{u}_i) &\forall \, i \in (0, 1, 2), \\
% \frac{\partial u_i}{\partial w} &= \D{\mathcal{T}}^{-1}(\imath \underline{n} 
%\hat{u}_i) &\forall \, i \in (0, 1, 2).
%\end{align}
%Dealiasing is implemented using the 2/3-rule.
%
%We finally note that the right hand sides of all the linear systems that need 
%to solved can be assembled fast using tailored matrix vector products taking 
%advantage of the fact that there are at most three distinct diagonals in any 
%matrix. This concludes the implementation required to solve the Navier-Stokes 
%equations by the incremental pressure correction method and the spectral 
%Shen-Fourier Galerkin method. We note that all terms in assembly and solve may 
%be performed using fast routines, at worst scaling like $O(N_x \log_2 N_x)$. 
%
%\section{Serial scaling}
%
%\section{Parallel scaling}
%
%
%

\section{Appendices}
%\subsection*{Fast Shen transforms}
%The fast forward and inverse Chebyshev transforms are used by the Shen 
%transforms. The code below computes both scalar products and complete 
%transforms for both 
%
%\begin{python}
%from numpy import *
%from scipy.fftpack import dct
%
%def fct(f, transform="GL"):
%    """Fast Chebyshev transform."""
%    N = f.shape[0]
%    f_hat = zeros(N)
%    if transform == "GL":
%        fk[:] = dct(fj, type=2, axis=0)            
%        fk /= N
%        fk[0] /= 2
%                
%    elif transform == "GC":
%        fk[:] = dct(fj, type=1, axis=0)/(N-1)
%        fk[0] /= 2
%        fk[-1] /= 2
%            
%    return fk
%
%def ifct(fk, transform="GL"):
%    """Inverse fast Chebyshev transform."""
%    if transform == "GL":
%        f = 0.5*dct(fk, type=3, axis=0)
%        f += 0.5*fk[0]
%    
%    elif transform == "GC":
%        f = 0.5*dct(fk, type=1, axis=0)
%        f += 0.5*fk[0]
%        f[::2] += 0.5*fk[-1]
%        f[1::2] -= 0.5*fk[-1]
%
%    return f
%
%def fastChebScalar(f, transform="GL"):
%    """Fast Chebyshev scalar product."""
%    N = f.shape[0]
%    if transform == "GL":
%        fk = dct(fj, type=2, axis=0)*pi/(2*N)
%        
%    elif transform == "GC":
%        fk = dct(fj, type=1, axis=0)*pi/(2*(N-1))
%    return fk
%
%def fastShenScalar(f, transform="GL"):
%    """Fast Shen scalar product"""                
%    fk = fastChebScalar(f, transform)
%    fk[:-2] -= fk[2:]
%    return fk
%
%def ifst(fk):
%    """Fast inverse Shen transform
%    """
%    N = len(fj)
%    wk = np.zeros(N, dtype=fk.dtype)        
%    wk[:-2] = fk[:-2] 
%    wk[2:] -= fk[:-2] 
%    return ifct(wk)
%    
%
%def fst(fj, transform="GL"):
%    """Fast Shen transform"""
%    fk = fastShenScalar(fj, transform)
%    
%    N = fj.shape[0]
%    if transform == "GL":
%        ck = np.ones(N-2); ck[0] = 2
%        
%    elif transform == "GC":
%        ck = np.ones(N-2); ck[0] = 2; ck[-1] = 2
%        
%    a = np.ones(N-4)*(-np.pi/2)
%    b = np.pi/2*(ck+1)
%    c = a.copy()
%    fk[:-2] = TDMA_1D(a, b, c, fk[:-2])
%        
%    return fk
%
%def fastShenScalar(f, transform="GL"):
%    """Fast Shen scalar product Neumann basis"""        
%    k  = wavenumbers(f.shape)
%    fk = fastChebScalar(f, transform)
%    fk[:-2] -= ((k/(k+2))**2) * fk[2:]
%    return fk
%
%def ifst(fk):
%    """Fast inverse Shen scalar transform Neumann basis"""
%    k = self.wavenumbers(fk.shape[0])
%    wk = np.zeros(fk.shape[0])    
%    wk[1:-2] = fk[1:-2]
%    wk[3:] -= (k[1:]/(k[1:]+2))**2*fk[1:-2]
%    f = ifct(wk, transform)
%    return f
%        
%def fst(f, transform="GL"):
%    """Fast Shen transform Neumann basis"""
%    fk = fastShenScalar(f, transform)
%    N = fj.shape[0]
%    k = wavenumbers(N)
%    ck = np.ones(N-3)
%    if transform == "GC": ck[-1] = 2
%    a = np.ones(N-5)*(-np.pi/2)*(k[1:-2]/(k[1:-2]+2))**2
%    b = np.pi/2*(1+ck*(k[1:]/(k[1:]+2))**4)
%    c = a.copy()
%    fk[1:-2] = TDMA_1D(a, b, c, fk[1:-2])
%    return fk
%
%# Cython implementation of TDMA    
%import numpy as np
%cimport cython
%cimport numpy as np
%#cython: boundscheck=False
%#cython: wraparound=False
%
%ctypedef fused T:
%    np.float64_t
%    np.complex128_t
%
%def TDMA_1D(np.ndarray[np.float64_t, ndim=1] a, 
%            np.ndarray[np.float64_t, ndim=1] b, 
%            np.ndarray[np.float64_t, ndim=1] c, 
%            np.ndarray[T, ndim=1] d):
%    cdef:
%        unsigned int n = b.shape[0]
%        unsigned int m = a.shape[0]
%        unsigned int k = n - m
%        int i
%        
%    for i in range(m):
%        d[i + k] -= d[i] * a[i] / b[i]
%        b[i + k] -= c[i] * a[i] / b[i]
%    for i in range(m - 1, -1, -1):
%        d[i] -= d[i + k] * c[i] / b[i + k]
%    for i in range(n):
%        d[i] /= b[i]
%        
%    return d
%    
%
%\end{python}
\subsection*{Matrices}

\begin{equation}
 \{\D{A}_{kj}\}_{k,j=0}^{N_x-2} = -\left(\D{\phi}^{''}_{j}, \D{\phi}_{k} \right)_w = \begin{cases}
 2\pi(k+1)(k+2), &j=k,\\
 4\pi(k+1), & j=k+2, k+4, k+6, \ldots, \\
 0, &\text{otherwise}.
 \end{cases}
\end{equation}

\begin{equation}
 \{\N{A}_{kj}\}_{k,j=1}^{N_x-2} = -\left(\N{\phi}^{''}_{j}, \N{\phi}_{k} \right)_w = \begin{cases}
 2\pi j^2(k+1)/(k+2), &j=k,\\
 4\pi j^2(k+1)/(k+2)^2, & j=k+2, k+4, k+6, \ldots, \\
 0, &j>k, \text{ or } k+j \text{ odd}.
 \end{cases}
\end{equation}

\begin{equation}
 \{\N{C}_{kj}\}_{k,j=0}^{N_x-4, N_x-2} = \left(\D{\phi}^{'}_j, \N{\phi}_k 
 \right)_w = \begin{cases}
 -\pi(k+1), &j=k-1,\\
 2\pi(k+1), & j=k+1, \\
 -\pi(k+1), & j=k+3, \\
 0, &\text{otherwise}.
 \end{cases}
\end{equation}

\begin{equation}
 \{\D{C}_{kj}\}_{k,j=0}^{N_x-2, N_x-4} = \left(\N{\phi}^{'}_j, \D{\phi}_k 
 \right)_w = \begin{cases}
 \pi (k-2)(k+1)/k, &j=k-3,\\
 -2 \pi(k+1)^2/(k+2), & j=k-1, \\
 \pi(k+1), & j=k+1, \\
 0, &\text{otherwise}.
 \end{cases}
\end{equation}

\begin{equation}
 \{\N{B}_{kj}\}_{k,j=0}^{N_x-4} = \{\N{B}_{jk}\}_{j,k=0}^{N_x-4} = (\N{\phi}_j, 
 \N{\phi}_k)_w = \begin{cases}
 \frac{\pi}{2} \left(c_k + 4 \left(\frac{k+2}{k+3} \right)^2 + c_{k+4} 
 \left(\frac{k+1}{k+3}\right)^2    \right), &j=k,\\
 -\pi \left( \frac{k+2}{k+3} + \frac{(k+4)(k+1)}{(k+5)(k+3)} \right), &j=k+2,\\
 \frac{\pi}{2} \frac{k+1}{k+3} , & j=k+4, \\
 0, &\text{otherwise}.
 \end{cases}
\end{equation}

\begin{equation}
 \{B_{kj}\}_{k,j=0}^{N_x} = (T_k, T_j)_w = \frac{c_k \pi}{2} \delta_{kj}
\end{equation}

For Gauss-Lobatto points we have $c_0=c_{N_x}=2$ and $c_k=1$ for $0<k<N_x$. For Gauss-Chebyshev points we have $c_0=2$ and $c_k=1$ for $k>0$. Note that for an $L^2_w$ scalar product we would have $c_{N_x}=1$ as well, but for the $l^2_w$ scalar product used here we get $c_{N_x}=2$. This disagreement, that follows from inexact quadrature at the highest mode, is not reported by Shen, but is reported by, e.g., Kopriva.

\end{document}