### 2016-09-20 paper.tex
Some speed results

\documentclass[11pt, oneside]{article}
\usepackage[utf8]{inputenc}
\usepackage{bm}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{ptex2tex, minted}
\usepackage{xcolor}
\usepackage{cancel}
\usepackage{caption}
\usepackage{subfigure}
\usepackage{algorithm}
\usepackage{algpseudocode}

\subfiglabelskip=0pt

\definecolor{gray}{gray}{0.97}
\colorlet{commentcolour}{green!50!black}
\colorlet{stringcolour}{red!60!black}
\colorlet{keywordcolour}{magenta!90!black}
\colorlet{exceptioncolour}{yellow!50!red}
\colorlet{commandcolour}{blue!60!black}
\colorlet{numpycolour}{blue!60!green}
\colorlet{literatecolour}{magenta!90!black}
\colorlet{promptcolour}{green!50!black}
\colorlet{specmethodcolour}{violet}
\colorlet{indendifiercolour}{green!70!white}

%\newcommand{\codetitlestyle}[1]{\small\textit{#1}\hspace{0.1cm}}
\newcommand{\belowtitleskip}{2pt}%\smallskipamount}
%\newcommand{\captionposition}{t}

%\newcommand{\mmo}[1]{\emph{#1}}

%\renewcommand{\ttdefault}{pcr}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[top=3cm,bottom=4cm,left=3cm,right=3.2cm,asymmetric]{geometry}

\lstset{
numbers=none,
aboveskip=1ex,
belowskip=1ex,
basicstyle=\ttfamily\footnotesize,
}

\lstdefinestyle{pythonstyle}{
%%\lstset{
%%keepspaces=true,
language=python,
showtabs=true,
tab=,
tabsize=2,
basicstyle=\ttfamily\footnotesize,%\setstretch{.5},
stringstyle=\color{stringcolour},
showstringspaces=false,
alsoletter={1234567890},
otherkeywords={\ , \}, \{, \%, \&, \|},
keywordstyle=\color{keywordcolour}\bfseries,
emph={and,break,class,continue,def,yield,del,elif ,else,%
except,exec,finally,for,from,global,if,import,in,%
lambda,not,or,pass,print,raise,return,try,while,assert},
emphstyle=\color{blue}\bfseries,
emph={[2]True, False, None},
emphstyle=[2]\color{keywordcolour},
emph={[3]object,type,isinstance,copy,deepcopy,zip,enumerate,reversed,list,len,dict,tuple,xrange,append,execfile,real,imag,reduce,str,repr},
emphstyle=[3]\color{commandcolour},
emph={Exception,NameError,IndexError,SyntaxError,TypeError,ValueError,OverflowError,ZeroDivisionError},
emphstyle=\color{exceptioncolour}\bfseries,
%upquote=true,
morestring=[s]{"""}{"""},
morestring=[s]{'''}{'''},
commentstyle=\color{commentcolour}\slshape,
emph={[4]1, 2, 3, 4, 5, 6, 7, 8, 9, 0,ode, fsolve, sqrt, exp, sin, cos, arccos, pi, array, norm, solve,float,complex, dot, arange, isscalar, max, sum, flatten, shape, reshape, find, any, all, abs, plot, linspace, legend, quad, polyval,polyfit, hstack,vector, concatenate,vstack,column_stack,empty,zeros,ones,rand,vander,grid,pcolor,eig,eigs,eigvals,svd,qr,tan,det,logspace,roll,min,mean,cumsum,cumprod,diff,vectorize,lstsq,cla,eye,xlabel,ylabel,squeeze},
emphstyle=[4]\color{commandcolour},
emph={[5]__init__,__add__,__mul__,__div__,__sub__,__call__,__getitem__,__setitem__,__eq__,__ne__,__nonzero__,__rmul__,__radd__,__repr__,__str__,__get__,__truediv__,__pow__,__name__,__future__,__all__},
emphstyle=[5]\color{specmethodcolour},
emph={[6]assert,range,yield},
emphstyle=[6]\color{keywordcolour}\bfseries,
emph={[7]def, return, and, print},
emphstyle=[7]\color{black}\bfseries,
% emph={[7]self},
% emphstyle=[7]\bfseries,
literate=*%
%{:}{{\literatecolour:}}{1}%
%{=}{{\literatecolour=}}{1}%
%{-}{{\literatecolour-}}{1}%
%{+}{{\literatecolour+}}{1}%
%{*}{{\literatecolour*}}{1}%
{/}{{\literatecolour/}}{1}%
{!}{{\literatecolour!}}{1}%
%{(}{{\literatecolour(}}{1}%
%{)}{{\literatecolour)}}{1}%
%{[}{{\literatecolour[}}{1}%
%{]}{{\literatecolour]}}{1}%
{<}{{\literatecolour<}}{1}%
{>}{{\literatecolour>}}{1}%
{>>>}{{\textcolor{promptcolour}{>>>}}}{1}%
,%
breaklines=true,
breakatwhitespace= true,
xleftmargin=\framemargin,
xrightmargin=\framemargin,
aboveskip=1ex,
belowskip=1ex,
frame=trbl, %trbl
numbers=none,
%frameround=tttt,
rulecolor=\color{black!40},
%framexleftmargin=\framemargin,
%framextopmargin=.1ex,
%framexbottommargin=.1ex,
%framexrightmargin=\framemargin,
%framexleftmargin=1mm, framextopmargin=1mm, frame=shadowbox, rulesepcolor=\color{blue},#1
%frame=tb,
%backgroundcolor=\color{yellow!10}
backgroundcolor=\color{gray}
%}
}

\newcommand{\N}[1]{\check{#1}}
\newcommand{\D}[1]{\overline{#1}}

\newcommand{\inpyth}{\lstinline[style=pythonstyle, basicstyle=\ttfamily]} %[]%

\lstnewenvironment{inpython}[1][]{
\lstset{style=pythonstyle, frame=trbl, belowcaptionskip=\belowtitleskip}
}{}

\newcommand{\includecode}[2][py]{\lstinputlisting[caption=#2,label=list:#2,style=pythonstyle,
float=!htpb]{#2}}

\newcommand{\mmo}[1]{({\bf mmo comment:} \emph{#1})}
\bibliographystyle{plain}

\title{Assessment of a Shen-Fourier turbulent channel flow solver for large-scale simulations}
\author{Mikael Mortensen and Diako Darian}
%\date{}							% Activate to display a given date or no date

\begin{document}
\maketitle
\section*{Abstract}

\section{Introduction}

\section{Preliminaries}
The Navier-Stokes equations used to describe turbulent flow in a doubly 
periodic channel can be written in rotational form as
\begin{align}
 \frac{\partial \bm{u}}{\partial t}   &= \bm{\mathcal{H}} + \nu 
 \nabla^2 \bm{u} - \nabla{P}, \notag \\
 \nabla \cdot \bm{u} &= 0, \label{eq:NS}
\end{align}
where $\bm{u}(\bm{x}, t)=(u, v, w)$ is the velocity vector, $\bm{x}=(x, y, z)$ 
and $t$ are position and time and $\bm{\mathcal{H}}(\bm{x}, t) = \bm{u}\times 
\bm{\omega}$ (where $\bm{\omega} = \nabla \times \bm{u}$)  represents the 
nonlinear convection.  The constant dynamic viscosity is denoted as $\nu$ and 
$P$ is a pressure modified to account for the kinetic energy, i.e., $P = p + 
\bm{u} \cdot \bm{u}/2$, where $p$ the instantaneous pressure 
normalized by a constant density. The computational domain is $\Omega=[-1, 
1]\times [0, L_y] \times [0, L_z]$ with channel walls located at $x=\pm 1$ such 
that no-slip applies at $ \bm{u}(\pm 1, y, z, t) = 0$. The walls are spanning 
the $y-z$ plane and the equations are periodic in the $y$ and $z$ directions 
with periodic lengths $L_y$ and $L_z$, respectively.

The Navier Stokes equations are discretized using Fourier basis functions for the periodic directions, and a combination of Chebyshev polynomials in the wall normal direction. Three different sets of basis functions and function spaces are relevant for the wall normal direction
\begin{align}
&  \phi_k(x) = T_k(x), & V_{N_x} &= \text{span}\{\phi_0, \phi_1,\ldots, \phi_{N_x}\} \label{eq:Tk}\\
& \D{\phi}_k(x) = T_k(x) - T_{k+2}(x), & \D{V}_{N_x} &= \text{span} \{ \D{\phi}_0, \D{\phi}_1, \ldots, \D{\phi}_{N_x-2} \} \label{eq:phiD}\\
& \N{\phi}_k(x) = T_k(x) - \frac{2(k+2)}{k+3} T_{k+2}(x) + 
\frac{k+1}{k+3} T_{k+4}, & \N{V}_{N_x} &= \text{span} \{\N{\phi}_0, \N{\phi}_1, 
\ldots, \N{\phi}_{N_x-4} \} \label{eq:phiN} 
\end{align}
where $T_k(x)$ is the $k$'th degree Chebyshev polynomial of the first kind. The 
basis functions and function spaces in (\ref{eq:phiD}) and (\ref{eq:phiN}) were 
suggested by Shen, and satisfy, respectively, the boundary conditions 
$\D{\phi}_k(\pm 1) = 0$, $\N{\phi}_k(\pm 1)=0$ and $\N{\phi}'_k(\pm 1)=0$. The 
last two function spaces may alternatively be written as $\D{V}_{N_x} = \{v \in 
V_{N_x}: v(\pm 1)=0 \}$ and $\N{V}_{N_x} = \{v \in V_{N_x}: v(\pm 1) = v'(\pm 
1) = 0 \}$.

Three-dimensional basis functions and function spaces that are periodic in $y$ 
and $z$ directions, can now be defined as
\begin{align}
  \psi_{\bm{k}}(\bm{x}) = \phi_{l}(x)e^{ \imath(\underline{m} y + \underline{n} z)}, \quad V_N &= \text{span} \{ \psi_{\bm{k}}(\bm{x}):\, \bm{k} \in K_N  \}, \\
  \D{\psi}_{\bm{k}}(\bm{x}) = \D{\phi}_{l}(x)e^{ \imath(\underline{m} y + \underline{n} z)}, \quad \D{V}_N &= \text{span} \{ \D{\psi}_{\bm{k}}(\bm{x}):\, \bm{k} \in \D{K}_N  \}, \\
  \N{\psi}_{\bm{k}}(\bm{x}) = \N{\phi}_{l}(x)e^{ \imath(\underline{m} y + \underline{n} z)}, \quad \N{V}_N &= \text{span} \{ \N{\psi}_{\bm{k}}(\bm{x}):\, \bm{k} \in \N{K}_N  \},
\end{align}
where $\imath=\sqrt{-1}$ and 
\begin{align}
K_N = \Big\{\bm{k} \in \mathcal{Z}\times\mathcal{R}^2 / &(l, \underline{m}, \underline{n}) \in \left( l, \frac{2 \pi m}{L_y}, \frac{2 \pi n}{L_z} \right), \text{where} \notag \\
 &(l, m, n) \in  [0, \ldots, N_x] \times [-\frac{N_y}{2},\ldots, 
 \frac{N_y}{2}-1] \times[-\frac{N_z}{2},\ldots, \frac{N_z}{2}-1] \Big\} 
 \label{eq:wavenumbermesh}
\end{align}
is the wavenumber mesh for $V_N$. The other two wavenumber meshes $\D{K}_N$ and 
$\N{K}_N$ are both equal to $K_N$ except from in index $l$ that 
ranges $[0, 1, \ldots, N_x-2]$ and $[0, 1, \ldots, N_x-4]$ for $\D{K}_N$ 
and $\N{K}_N$ respectively, see (\ref{eq:phiD}) and (\ref{eq:phiN}). We use 
notation $K_N^p(m, n)=K_N(0, m, n)$ when referring to only the two periodic 
dimensions of $K_N$.

The computational mesh in real space can be created using $N = (N_x, N_y, 
N_z)$ intervals, where the two periodic directions use uniform intervals. The 
computational mesh is given as
\begin{align}
  X_N = \Big\{ \bm{x} \in \mathcal{R}^3/&(x_i, y_j, z_k) \in \left( h(i), \frac{jL_y}{N_y}, \frac{kL_z}{N_z} \right), \text{where} \notag \\
  &(i, j, k) \in [0, 1, \ldots, N_x] \times [0, 1, \ldots, N_y-1] \times [0, 1, \ldots, N_z-1] \Big\} \label{eq:Xn}
\end{align}
where $x_i = h(i)$ represents either
\begin{equation}
 h(i) = \begin{cases}
   \cos \left(\frac{i \pi }{N_x} \right) \, &\forall \, i=0,1, \ldots, N_x \quad  \text{for Gauss-Lobatto points}, \\
   \cos \left(\frac{(2i +1)\pi}{2N_x+2} \right) \, &\forall \, i=0,1, \ldots, N_x \quad  \text{for Gauss-Chebyshev points}. \\
 \end{cases}
\end{equation}

For the Navier Stokes equations we look for solutions of the velocity 
components of the form
\begin{align}
u(\bm{x}, t) &= \frac{1}{N_yN_z}\sum_{\bm{k} \in \N{K}_N} \hat{u}_{\bm{k}}(t) 
\N{\psi}_{\bm{k}}(\bm{x}), \label{eq:u_solx} \\
v(\bm{x}, t) &= \frac{1}{N_yN_z}\sum_{\bm{k} \in \D{K}_N} \hat{v}_{\bm{k}}(t) 
\D{\psi}_{\bm{k}}(\bm{x}), \label{eq:u_soly} \\
w(\bm{x}, t) &= \frac{1}{N_yN_z}\sum_{\bm{k} \in \D{K}_N} \hat{w}_{\bm{k}}(t) 
\D{\psi}_{\bm{k}}(\bm{x}), \label{eq:u_solz}
\end{align}
where $\hat{u}_{\bm{k}}(t) = \hat{u}(l, {m}, {n}, t)$ are the expansion 
coefficients for the velocity component in $x$-direction (and similar for the 
other two components) and the scaling by $N_y$ and $N_z$ is merely for 
convenience 
and compliance with the definition used for the inverse discrete Fourier 
transform. Note that from now on we will simply use the notation $\hat{u}$ for 
$\hat{u}_{\bm{k}}(t)$, when it is possible to simplify without loss of clarity. 
Likewise we will simply use $u$ for $u(\bm{x}, t)$.

If the expansion coefficients $\hat{u}$  are known for the entire wavenumber 
mesh (\ref{eq:wavenumbermesh}), then the expression (\ref{eq:u_solx}) may be 
evaluated on the mesh (\ref{eq:Xn}) using fast Fourier and fast Shen inverse 
transforms as
\begin{align}
u(x_i, y_j, z_k, t) &= \underbrace{\frac{1}{N_z}\sum_n 
\underbrace{\frac{1}{N_y} \sum_m \underbrace{\sum_l 
\hat{u}(l,m,n,t)\N{\phi}_l(x_i)}_{\N{\mathcal{S}}_x^{-1}} e^{\imath 
\underline{m} 
y_j}}_{\mathcal{F}_y^{-1}} e^{\imath \underline{n} z_k}}_{\mathcal{F}_z^{-1}} , 
\quad \forall \, \bm{x} \in X_N, \notag \\
  u &= \N{\mathcal{T}}^{-1}(\hat{u}) =  \mathcal{F}^{-1}_z (\mathcal{F}^{-1}_y 
  (\N{\mathcal{S}}^{-1}_x (\hat{u}))), \quad \forall \, \bm{x} \in X_N.  
  \label{eq:ifft} 
\end{align}
Here $\N{\mathcal{T}}^{-1}$ is used as short notation for the complete inverse 
transform, $\mathcal{F}_{y}^{-1}$ and $\mathcal{F}_{z}^{-1}$ represent inverse 
Fourier transforms along directions $y$ and $z$ respectively, and the inverse 
Shen transform $\N{\mathcal{S}}_{x}^{-1}$ is performed along the wall normal 
$x$ direction. Note that the transforms are applied sequentially to the entire 
computational mesh. The inverse Shen transform may be computed using fast 
Chebyshev transforms for all the three bases in (\ref{eq:Tk}, \ref{eq:phiD}, 
\ref{eq:phiN}). The inverse transforms are slightly different for the three 
bases and 
${\mathcal{S}}^{-1}, \D{\mathcal{S}}^{-1}$ and $\N{\mathcal{S}}^{-1}$ are used 
to distinguish between them with obvious notation.

A three-dimensional scalar product in the weighted $[L^2_w(\Omega)]^3$ space is defined as
\begin{align}
 \left<u, v\right>_w &= \int_{\Omega} {u(\bm{x}) v^*(\bm{x})}w\,dxdydz, 
\end{align}
where $v^*$ is the comlex conjugate of $v$ and the weights $w$ are unity for 
periodic directions and  $w=1/\sqrt{1-x^2}$ for the inhomogeneous direction. In 
this work we will make use of the discrete weighted $[l^2_w(\Omega)]^3$ space, 
where quadrature is employed for integration. As such the scalar product 
corresponds to sums that may be computed using fast transforms. Using 
$\N{\psi}_{\bm{k}}$ for $v$ we obtain
\begin{align}
 \left< u, \N{\psi}_{\bm{k}} \right>_w &= h \underbrace{\sum_i 
 \underbrace{\sum_j \underbrace{\sum_k u(x_i, y_j, z_k, t)  e^{-\imath 
 \underline{n} z_k}}_{\mathcal{F}_n}  e^{-\imath \underline{m} y_j} 
 }_{\mathcal{F}_m} \N{\phi}_{l}(x_i) w_i}_{\N{\mathcal{S}}_l},   \notag \\
  &=  h \N{\mathcal{S}}(u) = h \N{\mathcal{S}}_{l} (\mathcal{F}_{{m}} 
  (\mathcal{F}_{{n}}(u))), \quad \forall \, \bm{k} \in \N{K}_N. \label{eq:sst1}
\end{align}
Here $h\N{\mathcal{S}}$ denotes the complete three dimensional scalar product 
and $h = L_yL_zN_y^{-1}N_z^{-1}$ is a constant. $\mathcal{F}_{{n}}$ and 
$\mathcal{F}_{{m}}$ represent discrete Fourier transforms in $z$- and 
$y$-directions, respectively, and $\N{\mathcal{S}}_{l}(\cdot) = (\cdot, 
\N{\phi}_l)_w$ is used to represent the Shen scalar product in the 
$x$-direction. The weights, $w_i$, required for the Shen transforms are 
defined as
\begin{equation}
 w_i = \begin{cases}
       \frac{\pi}{c_i N_x} &\forall \, i=0,1,\ldots, N_x \quad  \text{for 
       Gauss-Lobatto points},\\
       \frac{\pi}{N_x+1} &\forall \, i=0,1,\ldots, N_x  \quad \text{for 
       Gauss-Chebyshev points},      
 \end{cases}
\end{equation}
where $c_0 = c_{N_x} = 2$ and $c_i = 1$ for $0 < i < N_x$.\footnote{Note that 
$c_{N_x}=1$ for the continuous scalar product $L^2_w(\Omega)$ (see, e.g., Eq. 
(2.7) of Shen2), but for the discrete $l^2_w(\Omega)$ it follows that 
$c_{N_x}=2$ due to inexact quadrature for the highest wavenumber.}

The scalar product  $h\N{\mathcal{S}}$ in (\ref{eq:sst1}) does not represent a 
complete transform. To find a transformation between physical $u$ and spectral 
$\hat{u}$ we make use of Eq. (\ref{eq:u_solx}) directly and use the discrete 
orthogonality of the Fourier basis functions
\begin{align}
%\left<u, \N{\psi}_{\bm{k}}\right>_w &= \frac{1}{N_yN_z}\left< 
%\sum_{\bm{m}=(q,r,s) \in \N{K}_N} \hat{u}_{\bm{m}} \N{\psi}_{\bm{m}}, 
%\N{\psi}_{\bm{k}} \right>_w, \notag \\
\left<u, \N{\psi}_{\bm{k}}\right>_w &= \frac{1}{N_yN_z}\left< 
\sum_{(q,r,s) \in \N{K}_N} \hat{u}(q,r,s,t) \N{\psi}(q,r,s), 
\N{\psi}(l,m,n,t) \right>_w, \notag \\
           &= h \sum_{q=0}^{N_x-4} \sum_{i=0}^{N_x} \N{\phi}_{q}(x_i) 
           \N{\phi}_{l}(x_i) w_i \, \hat{u}(q, {m}, {n}, t), \notag \\
           &= h \sum_{q=0}^{N_x-4} \N{B}_{lq} \hat{u}(q, {m}, {n}, t), 
           \label{eq:sst2}
\end{align}
where $\N{B}_{lq} = (\N{\phi}_q, \N{\phi}_l)_w = 
\sum_{i=0}^{N_x} \N{\phi}_{q}(x_i) \N{\phi}_{l}(x_i) w_i$ is a 
pentadiagonal mass matrix. The complete transformation is now obtained by 
setting Eq. (\ref{eq:sst2}) equal to (\ref{eq:sst1}) and solving for $\hat{u}$. 
Moving to matrix notation we get
\begin{align}
h\sum_{q=0}^{N_x-4} \N{B}_{lq}\, \hat{u}(q, {m}, {n}, t) &= 
h\N{\mathcal{S}}(u)(l, m, n, t), \notag \\
 \hat{u}_{\bm{k}}(t) &= \N{\mathcal{T}}(u) =  \N{B}^{-1}\N{\mathcal{S}}(u), 
 \quad \forall \, \bm{k} \in \N{K}_N, \label{eq:fstB}
\end{align}
where $\N{\mathcal{T}}(u)$ denotes the complete transformation, such that $u = 
\N{\mathcal{T}}^{-1}(\N{\mathcal{T}}(u))$. Note that since $\N{B}$ assembles to 
a pentadiagonal matrix the solution ($\N{B}^{-1}$) can be obtained very fast 
and 
the complete 
transformation thus requires a fast Chebyshev transform ($O(N_x \log N_x)$) and 
a fast linear algebraic solve ($O(N_x)$). Similar transforms $\mathcal{T}$ 
and $\D{\mathcal{T}}$ are defined for the two other bases (\ref{eq:Tk}) and 
(\ref{eq:phiD}), using mass matrices $B_{lq}=(T_q, T_l)_w$ (diagonal) and 
$\D{B}_{lq}=(\D{\phi}_l, \D{\phi}_q)_w$ (tridiagonal) and scalar products 
$\mathcal{S}_l(\cdot) = (\cdot, T_l)_w$ and $\D{\mathcal{S}}_l(\cdot) = (\cdot, 
\D{\phi}_l)_w$.

\section{Discretization of Navier Stokes equations}
\label{sec:discretizationNS}
The Navier Stokes equations (\ref{eq:NS}) are solved using a scheme 
proposed by Kim, Moin and Moser (1987). This scheme is developed by taking the 
Laplacian of the wall normal momentum equation and the curl of the momentum 
equation. Following elimination of the pressure, the equations to solve are 
\begin{align}
\frac{\partial}{\partial t} \nabla^2 u &= h_u + \nu \nabla^4 u, 
\label{eq:biharmonic} \\
\frac{\partial g}{\partial t} &= h_g + \nu \nabla^2 g, \label{eq:g} \\
f + \frac{\partial u}{\partial x} &= 0, \label{eq:f}
\end{align}
where
\begin{align}
f &= \frac{\partial v}{\partial y} + \frac{\partial w}{\partial z}, \\
g &= \frac{\partial w}{\partial y} - \frac{\partial v}{\partial z}, \\
h_u = -\frac{\partial}{\partial x} &\left( \frac{\partial 
\mathcal{H}_y}{\partial y} + \frac{\partial \mathcal{H}_z}{\partial z} \right) 
+ \left(\frac{\partial^2}{\partial y^2} + \frac{\partial^2}{\partial z^2} 
\right) \mathcal{H}_x ,
\\
h_g &= \frac{\partial \mathcal{H}_z}{\partial y} - \frac{\partial 
\mathcal{H}_y}{\partial z}.
\end{align}
The two remaining velocity components are computed from the definitions of $f$ 
and $g$. The biharmonic equation (\ref{eq:biharmonic}) is solved with boundary 
conditions $u(\pm 1) = u'(\pm 1) = 0$, where the Neumann conditions follow 
from the continuity equation. The boundary conditions for Eq. (\ref{eq:g}) are 
$g(\pm 1) = 0$. The variational formulation becomes: Find ${u} \in 
\N{V}_N$, ${g} \in \D{V}_N$ and ${f} \in \D{V}_N$ such that
\begin{align}
	\frac{\partial }{\partial t} \left< \nabla^2 u, \N{\psi}\right>_w &= 
	\left<h_u, \N{\psi} \right>_w + \nu \left<\nabla^4u, \N{\psi}\right>_w 
	&\forall \N{\psi} \in \N{V}_N, \label{eq:u1} \\
	\frac{\partial}{\partial t}\left<g, \D{\psi}\right>_w &= \left<h_g, 
	\D{\psi}\right>_w + \nu 
	\left<\nabla^2 g, \D{\psi}\right>_w &\forall \D{\psi} \in \D{V}_N ,
	\label{eq:g1} \\
	\left<f, \D{\psi}\right>_w &= \left<\frac{\partial u}{\partial x}, 
	\D{\psi}\right>_w &\forall \D{\psi} \in \D{V}_N. \label{eq:f1}
\end{align}
The two remaining velocity components are computed by projection to the 
Dirichlet space $\D{V}_N$: Find 
${v} \in \D{V}_N$ and $w \in \D{V}_N$ such that
\begin{align}
\left<f, \D{\psi}\right>_w &= \left<\frac{\partial v}{\partial y} + 
\frac{\partial w}{\partial z}, \D{\psi}\right>_w \, &\forall \D{\psi} \in 
\D{V}_N, \label{eq:f2} \\
\left<g, \D{\psi}\right>_w &= \left<\frac{\partial w}{\partial y}  - 
\frac{\partial v}{\partial z}, \D{\psi}\right>_w \, &\forall \D{\psi} \in 
\D{V}_N, \label{eq:g2}
\end{align}
which simplifies considerably because all the derivatives are in periodic 
directions. Written in spectral space Eqs. (\ref{eq:f2}) and (\ref{eq:g2}) 
become simply
\begin{align}
\hat{f}_{\bm{k}} &= \imath \underline{m}\, \hat{v}_{\bm{k}} + \imath 
\underline{n}\, \hat{w}_{\bm{k}} &\forall \bm{k} \in \D{K}_N^0, \label{eq:f3} \\
\hat{g}_{\bm{k}} &= \imath \underline{m}\, \hat{w}_{\bm{k}} - \imath 
\underline{n}\, \hat{v}_{\bm{k}} & \forall \bm{k} \in \D{K}_N^0, \label{eq:g3}
\end{align}
where $\D{K}_N^0$ is used to denote that these equations can be solved for all 
wavenumbers except $m=n=0$. For $m=n=0$ we solve instead the 
corresponding transformed momentum equations in $y$ and $z$ directions:
\begin{align}
\frac{\partial }{\partial t} \D{B}_{lq}\hat{v}(q, 0, 0, t) &= 
\D{B}_{lq}\hat{\mathcal{H}}_y(q, 0, 0, t) + \nu \D{A}_{lq} \hat{v}(q, 0, 0, t) 
& \forall\, l \in [0, 1, 
\ldots, N_x-2], \label{eq:v0}\\
\frac{\partial }{\partial t} \D{B}_{lq}\hat{w}(q, 0, 0, t) &= 
\D{B}_{lq}\hat{\mathcal{H}}_z(q, 0, 0, t) + \nu \D{A}_{lq} \hat{w}(q, 0, 0, t) 
& \forall\, l \in [0, 1, \ldots, N_x-2].\label{eq:w0}
\end{align}

Equations (\ref{eq:u1}-\ref{eq:f1}) are also solved in spectral space. 
Inserting for (\ref{eq:u_solx}) and a similar expansion for $g$ in $\D{V}_N$, 
the inner products required to solve Eqs. (\ref{eq:u1}-\ref{eq:f1}) are
\begin{align}
\left<\nabla^4u, \N{\psi}_{\bm{k}}\right>_w &= h\left[ \left( 
\N{\phi}_q^{''''}, 
\N{\phi}_l\right)_w -2(\underline{m}^2+\underline{n}^2) \left( \N{\phi}_q^{''}, 
\N{\phi}_l\right)_w + (\underline{m}^2+\underline{n}^2)^2\left( \N{\phi}_q, 
\N{\phi}_l\right)_w  \right] \hat{u}_q, \\
\left< \nabla^2 u, \N{\psi}_{\bm{k}}\right>_w &= h\left[\left( \N{\phi}_q^{''}, 
\N{\phi}_l\right)_w - (\underline{m}^2+\underline{n}^2)\left( \N{\phi}_q, 
\N{\phi}_l \right)_w \right] \hat{u}_q, \\
\left< \nabla^2 g, \D{\psi}_{\bm{k}}\right>_w &= h\left[\left( \D{\phi}_q^{''}, 
\D{\phi}_l\right)_w - (\underline{m}^2+\underline{n}^2)\left( \D{\phi}_q, 
\D{\phi}_l \right)_w \right] \hat{g}_q, \\
\left<\frac{\partial u}{\partial x}, \D{\psi}_{\bm{k}}\right>_w &=
h\left(\N{\phi}_q^{'}, \D{\phi}_l\right)_w \hat{u}_q, \\
\left<h_u, \N{\psi}_{\bm{k}} \right>_w &= 
h\left[-(\underline{m}^2+\underline{n}^2) 
\N{\mathcal{S}}(\mathcal{H}_x) - \imath \underline{m}\left(\D{\phi}_q^{'}, 
\N{\phi}_l \right)_w \hat{\mathcal{H}}_{y, q} - \imath 
\underline{n}\left(\D{\phi}_q^{'}, \N{\phi}_l \right)_w 
\hat{\mathcal{H}}_{z, q}\right], 
\\
\left< h_g, \D{\psi}_{\bm{k}} \right>_w &= h\left[ \imath \underline{m}\, \D{B} 
\hat{\mathcal{H}}_z - \imath \underline{n}\, \D{B} \hat{\mathcal{H}}_{y} 
\right],
\end{align}
where for brevity in notation (as before) it is understood that the matrices 
act along the first dimension of the transformed variables, i.e.,  $ 
\left(\N{\phi}_q^{'}, \D{\phi}_l\right)_w \hat{u}_q$ is short for the matrix 
vector product $ \sum_{q=0}^{N_x-4}\left( \N{\phi}_q^{'}, \D{\phi}_l\right)_w 
\hat{u}(q, {m}, {n}, t)$. The inner products are used to set up linear systems 
of equations for the inhomogeneous wall normal direction. All inner products 
$\left(\cdot, \cdot\right)_w$ can be represented by sparse matrices.
%\begin{align}
%\frac{\partial }{\partial t} \left( \N{\psi}_j^{''}, \N{\psi}_k\right)_w 
%\hat{u}_j &=  \left(h_u, \N{\psi}_k \right)_w + \nu \left(\N{\psi}^{''''}_j, 
%\N{\psi}_k\right)_w \hat{u}_j, \, &\forall \N{\psi}_k \in \N{V}_N, \\
%\frac{\partial}{\partial t}\left(\D{\psi}_j, \D{\psi}_k\right)_w \hat{g}_j &= 
%\left(h_g, \D{\psi}_k\right)_w + \nu \left(\D{\psi}_j^{''}, 
%\D{\psi}_k\right)_w 
%\hat{g}_j, \, &\forall \D{\psi}_k \in \D{V}_N \\
%\left(\D{\psi}_j, \D{\psi}_k\right)_w \hat{f}_j &= \left(\N{\psi}_j^{'}, 
%\D{\psi}_k\right)_w \hat{u}_j, \, &\forall \D{\psi}_k \in \D{V}_N.
%\end{align}
The mass matrices, stiffness matrices $\N{A}_{lq} = \left( \N{\phi}_q^{''}, 
\N{\phi}_l\right)_w, 
\D{A}_{lq} = \left( \D{\phi}_q^{''}, \D{\phi}_l\right)_w 
$ and $\N{Q}_{lq} = \left(\N{\phi}^{''''}_q, \N{\phi}_l\right)_w$ are given by 
Shen, whereas $\left(\N{\phi}_q^{'}, \D{\phi}_l\right)_w$ and 
$\left(\D{\phi}_q^{'}, \N{\phi}_l\right)_w$ are given in the 
appendix.

It remains to discretize the equations in time. Kim, Moin and Mansour use 
Crank-Nicolson for the diffusive terms and Adams-Bashforth for the nonlinear 
convective terms, whereas a third order Runge-Kutta method is used for the 
nonlinear term by Moser, Kim and Mansour (1999). We have implemented both 
approaches, but for simplicity only the first one is described here. 
Discretization in time is performed using a constant time step $\triangle t$, 
such that the time is represented discretely as $t_n = n \triangle t, \, n \in 
\mathcal{Z}^{\ge}$. After 
discretizing in time, equations (\ref{eq:u1}) and (\ref{eq:g1}) are transformed 
to a matrix form that can be solved for $\hat{u}^{n+1}$ and $\hat{g}^{n+1}$ 
given wavenumbers $m$ and $n$
\begin{align}
\N{H}\hat{u}^{n+1} & = \left(2\N{A} - 2\underline{z}^2\N{B} - \N{H} 
\right)\hat{u}^{n} + \N{\mathcal{S}}(h_v^{n+1/2}) \triangle t & 
\forall \, {\bm{k}} \in \N{K}_N, 
\label{eq:ufin}\\ 
\D{H} \hat{g}^{n+1} &= \left(2 \D{B}-\D{H}\right) 
\hat{g}^{n} + \D{\mathcal{S}}(h_g^{n+1/2}) \triangle t &\forall \, 
{\bm{k}} \in 
\D{K}_N. \label{eq:gfin}
\end{align}
The coefficient matrices are given as
\begin{align}
\N{H}_{lq}(m, n) &= -\frac{\nu \triangle t}{2}\N{Q}_{lq} + \left( 1 + \nu 
\triangle t 
\underline{z}^2 \right) \N{A}_{lq} - \frac{2\underline{z}^2 + \nu \triangle t 
\underline{z}^4}{2} \N{B}_{lq} &\forall m, n \in K_N^p, 
\label{eq:Biharmonic_matrix} \\
\D{H}_{lq}(m, n) &= -\frac{\nu \triangle 
t}{2}\D{A}_{lq} + (1 + \frac{\nu \triangle t \underline{z}^2}{2}) \D{B}_{lq} 
&\forall m, n \in K_N^p, \label{eq:Helmholtz_matrix}
\end{align}
where $\underline{z}^2 = \underline{m}^2 + \underline{n}^2$. The convective 
terms are computed with a forward Adams-Bashforth projection, such that 
$\N{\mathcal{S}}(h_v^{n+1/2}) = 
\N{\mathcal{S}}(1.5h_v^{n} - 0.5 h_v^{n-1})$ and $\D{\mathcal{S}}(h_g^{n+1/2}) 
= \D{\mathcal{S}}(1.5h_g^{n} - 0.5 h_g^{n-1})$. Equations 
(\ref{eq:v0}) and (\ref{eq:w0}) are also discretized in time using 
Crank-Nicolson and Adams-Bashforth, and we obtain
\begin{align}
\left(-\frac{\nu \triangle t}{2}\D{A} + \D{B} \right)\hat{v}_0^{n+1} &= 
\left(\frac{\nu \triangle t}{2}\D{A} + \D{B} 
\right)\hat{v}^{n}_0 + \triangle t\D{B}\hat{\mathcal{H}}_{y,0}^{n+1/2} , 
\label{eq:v00} \\
\left(-\frac{\nu \triangle t}{2}\D{A} + \D{B} \right)\hat{w}_0^{n+1} &= 
\left(\frac{\nu \triangle t}{2}\D{A} + \D{B} 
\right)\hat{w}^{n}_0 + \triangle t\D{B}\hat{\mathcal{H}}_{z,0}^{n+1/2}, 
\label{eq:w00}
\end{align}
where we have used notation $\hat{v}^{n}_0 = \hat{v}(l, 0, 0, t_n)$, 
$\hat{\mathcal{H}}^{n}_{y,0} = \hat{\mathcal{H}}_y(l, 0, 0, t_n)$ and similar 
for $\hat{w}^n_0$ and $\hat{\mathcal{H}}^n_{z,0}$.

\section{Implementation}
\label{sec:implementation}
The solution procedure for the numerical method described in 
Sec.~\ref{sec:discretizationNS} is given in Algorithm \ref{alg:NS}

\begin{algorithm}
	\caption{Solution procedure for Navier Stokes equations}
    \label{alg:NS}
    \begin{algorithmic}[1]
    	\State \text{Initialize} $\bm{u}(\bm{x}, 0)$ and compute 
    	$\hat{g}_{\bm{k}}(0)$ and $\hat{u}_{\bm{k}}(0)$
		\State \text{Compute LU decomposition of} $\N{H}(m,n), \D{H}(m,n) 
		\,\forall \, m, n \in K_N^p$
		\State $t=0; n=0$
    	\While{$t<T$}
	    	\State \text{Compute nonlinear convection} $\bm{\mathcal{H}}^n$
	    	\State \text{Compute right hand sides of Eqs.~}(\ref{eq:ufin}), 
	    	(\ref{eq:gfin})
		    \State \text{Solve Eq.}(\ref{eq:ufin}) for 
		    $\hat{u}^{n+1}_{\bm{k}} \, \forall \,\bm{k} \in \N{K}_N$
		    \State \text{Solve Eqs.}(\ref{eq:gfin}) for 
		    $\hat{g}^{n+1}_{\bm{k}} \, \forall\, \bm{k} \in \D{K}_N$
		    \State \text{Solve Eqs.}(\ref{eq:f3}, \ref{eq:g3}) for  
		    $\hat{v}_{\bm{k}}, \hat{w}_{\bm{k}}\, \forall\, \bm{k} 
		    \in \D{K}_N^0$
		    \State \text{Solve Eqs.}(\ref{eq:v00}, \ref{eq:w00}) for 
		    $\hat{v}_0^{n+1}, \hat{w}_0^{n+1}\, 
		    \forall \, l \in [0, 1, \ldots, N_x-2] $
		    \State $n \gets n + 1$
		    \State $t \gets t+\triangle t$
		    \State \text{Update to new time step}
	    \EndWhile

    \end{algorithmic}
\end{algorithm}

The major computational cost comes from  computing the nonlinear convection 
and solving for Eqs.~(\ref{eq:ufin}, \ref{eq:gfin}). The nonlinear 
convection 
term $\bm{\hat{\mathcal{H}}}$ is computed using a pseudo-spectral method with 
the 3/2-rule for de-aliasing in the periodic directions (ref Canuto). The 
transforms are carried out using fast Fourier and fast Chebyshev transforms 
that are all of order $O(N\log N)$ per one-dimensional transform of size 
N. The Shen transforms make use of the Chebyshev transforms that are using 
discrete cosine transforms. The forward scalar product is implemented as shown 
in Algorithm \ref{alg:fst}. A complete transform is obtained through a scalar 
product followed by solving the linear system involving the mass 
matrix, just like shown in Eq. (\ref{eq:fstB}) for space $\N{V}_N$ and similar 
for $\D{V}_N$ and $V_N$. Since the mass matrices are diagonal, tridiagonal and 
pentadiagonal for $V_N, \D{V}_N$ and $\N{V}_N$ respectively, solving the linear 
system is $O(N)$ and the complete transform still scales like $O(N\log N)$. 
\begin{algorithm}
	\caption{Forward scalar product for all spaces $V_N, \D{V}_N, \N{V}_N$. 
	Here "dct" is the discrete cosine transform from SciPy.}
	\label{alg:fst}
	\begin{algorithmic}[1]
	\Procedure{fastScalar}{}
		\State \textbf{Input:} $\{f_j\}_{j=0}^{N}$
		\State \textbf{Input:} points
		\State \textbf{Input:} space
		\If{ points = "Gauss-Lobatto"}
			\State $\{ w_k\}_{k=0}^{N} \gets \text{dct}(f, \text{type=2}, 
			\text{axis=0})\frac{\pi}{2 N}$
		\Else{ points = "Gauss-Chebyshev"}
		    \State $\{ w_k\}_{k=0}^{N} \gets \text{dct}(f, \text{type=1}, 
		    \text{axis=0}) \frac{\pi}{2 (N-1)}$
	    \EndIf
	    \If{ space = $\D{V}_N$}
	        \For{$k=0$ to $N-2$}
		        \State $\hat{f}_k \gets w_k - w_{k+2}$
	        \EndFor
			\State \Return $\{\hat{f}_k\}_{k=0}^{N-2}$	        
	    \ElsIf{ space = $\N{V}_N$}	                
  	        \For{$k=0$ to $N-4$}
       	        \State $\hat{f}_k \gets w_k - \frac{2(k+2)}{k+3} w_{k+2} + 
       	        \frac{k+1}{k+3}w_{k+4}$
   	        \EndFor
   	        \State \Return $\{\hat{f}_k\}_{k=0}^{N-4}$
   	    \Else{}
   	        \For{$k = 0$ to $N$}
     	        \State $f_k \gets \frac{2}{\pi} w_k $
     	    \EndFor
     	    \State $f_0 \gets f_0 / 2$
     	    \If{ points = "Gauss-Lobatto"}
	     	\State $f_N \gets f_N / 2$	     	
     	    \EndIf
			\State \Return $\{f_k\}_{k=0}^{N}$    
	    \EndIf
	\EndProcedure
\end{algorithmic}
\end{algorithm}

Shen \cite{shen1995} states that it is possible to solve Eq.~(\ref{eq:gfin}) 
with 
essentially the same number of operations as a pentadiagonal solver. We note 
that the matrix 
$\D{H}$ is of type upper Hessenberg and a solution
procedure is simply to perform a direct LU decomposition (se, e.g., 
\cite{stewart98}), which leads to a 
lower triangular  
matrix $L$ with only one subdiagonal and an upper triangular matrix $U$ which 
is dense. However, each row in $U$ 
contains at most three distinct values at $U_{k,k}, U_{k,k+2}$ and $U_{k,k+4}$, 
and then $U_{k,j} = U_{k,k+4}\, \forall \, j = k+6, k+8, \ldots N$. 
Consequently 
only three diagonals in $U$ need storage and the back solve can be traversed 
very efficiently in $O(N)$. Note also that the system could be decoupled into 
odd and even coefficients, which leads to two subsystem that may be trivially 
solved 
simultaneously in two threads. For optimal 
performance, the odd and even coefficient would then need to be stored 
contiguously in computer memory, and not alternately, which would be the normal 
way of storing the coefficients.

%We will now describe very efficient direct solvers for Eqs.~(\ref{eq:ufin}, 
%\ref{eq:gfin}) that are of order $O(N)$. The solvers are generated using 
%clever 
%direct LU decomposition and only a few diagonals need to be stored for each 
%final lower (L) and upper (U) triangular matrix.

%To solve Eq. (\ref{eq:gfin}) we perform an LU decomposition of the coefficient 
%matrix $\D{H}$. $\D{H}$ consists of one subdiagonal, and every 
%second upper diagonal is zero. As such, we may split the matrix into two 
%smaller (and decoupled) matrices for odd and even coefficients $\D{H}^e_{k,j} 
%= 
%\D{H}_{2k,2j}$ and $\D{H}^o_{k,j} = \D{H}_{2k+1,2j+1}$, where the two matrices 
%$\D{H}^{e,o}$ are of type upper 
%Hessenberg, with only 4 distinct values on each row. This structure of $\D{H}$ 
%allows us to perform a very efficient tailored LU decomposition, as shown in 
%Algorithm (\ref{alg:lu}), and a very efficient tailored solve through 
%Algorithm (\ref{alg:lusolve}). The efficiency follows from the fact that 
%$U^{o,e}_{kj} = U^{o,e}_{kk+2} \, \forall \, j > k+2$, i.e., the upper 
%triangular matrix has a maximum of three distinct values for each row, which 
%allows for a very fast solve through back substitution. 
%Regarding Algorithm 
%(\ref{alg:lusolve}), we note that the ability to decouple the system into odd 
%and even coefficients leads to two subsystem that may be trivially solved 
%simultaneously in two threads by letting one thread call LUSolveoddeven with 
%even coefficients, and the second thread call it for odd. For optimal 
%performance, the odd and even coefficient would then need to be stored 
%contiguously in computer memory, and not alternately, which would be the 
%normal 
%way of storing the coefficients.

%\begin{algorithm}
%\caption{LU factorization of matrix $H_{k,j}$, that may be split into two 
%decoupled upper Hessenberg matrices $H^e_{k,j} = H_{2k,2j}$ and $H^o_{k,j} = 
%H_{2k+1,2j+1}$ for even and odd coefficients. Return LU factorization for even 
%and odd coefficients, such that $L^eU^e=H^e$ and $L^oU^o=H^o$. Note that 
%$L^{e}$ and $L^{o}$ each have one single nonzero diagonal in addition to the 
%unity main diagonal, and as such only a vector is stored. Each $U^{e}$ and 
%$U^{o}$ has three distinct diagonals and $U^{e}_{kj}=U^{e}_{k, k+2} \,\forall 
%j 
%> k+2, j< M^e$ and $U^{o}_{kj}=U^{o}_{k, k+2} \,\forall j > k+2, j< M^o$. As 
%such, a sparse three-diagonal storage may be used for $U^e$ and $U^o$.}
%\label{alg:lu}
%\begin{algorithmic}[1]
%\Procedure{LUoddeven}{}
%  \State \textbf{Input:} $\{H\}_{k,j=0}^{N-2}$
%  \State \textit{integer} $M^e \gets (N-2)/2$
%  \State \textit{integer} $M^o \gets (N-3)/2$
%  \State ${U}^e_{0j} \gets {H}_{0,2j}, \, \forall j=0,1,2 $
%  \State ${U}^o_{0j} \gets {H}_{1,2j+1}, \, \forall j=0,1,2 $
%  \For{$k=1, 2,  \ldots, M^e-2$}
%    \State ${L}^e_{k} \gets \frac{{H}_{2k,2(k-1)}}{{U}^e_{k-1,k-1}}$
%    \State ${U}^e_{kj} \gets {H}_{2k,2j} - {L}^e_{k}{U}^e_{k-1,j}, \, \forall 
%j=k, k+1, k+2$
%    \State ${L}^o_{k} \gets \frac{{H}_{2k+1,2k-1}}{{U}^o_{k-1,k-1}}$
%    \State ${U}^o_{kj} \gets {H}_{2k+1,2j+1} - {L}^o_{k}{U}^o_{k-1,j}, \, 
%\forall j=k, k+1, k+2$
%  \EndFor
%  
%  \State $k \gets M^e-1$
%  \State ${L}^e_{k} \gets \frac{{H}_{2k,2(k-1)}}{{U}^e_{k-1,k-1}}$
%  \State ${U}^e_{kj} \gets {H}_{2k,2j} - {L}^e_{k}{U}^e_{k-1,j}, \, \forall 
%j=k, k+1$
%  \State ${L}^o_{k} \gets \frac{{H}_{2k+1,2k-1}}{{U}^o_{k-1,k-1}}$
%  \State ${U}^o_{kj} \gets {H}_{2k+1,2j+1} - {L}^o_{k}{U}^o_{k-1,j}, \, 
%\forall j=k, k+1$
%  \State $k \gets M^e$
%  \State ${L}^e_{k} \gets \frac{{H}_{2k,2(k-1)}}{{U}^e_{k-1,k-1}}$
%  \State ${U}^e_{kk} \gets {H}_{2k,2k} - {L}^e_{k}{U}^e_{k-1,k}$
%  \If{$M^o == M^e$ }
%    \State ${L}^o_{k} \gets \frac{{H}_{2k+1,2k-1}}{{U}^o_{k-1,k-1}}$
%    \State ${U}^o_{kk} \gets {H}_{2k+1,2k+1} - {L}^o_{k}{U}^o_{k-1,k}$
%
%  \EndIf  
%  \State \Return $\{L^e_k\}_{k=0}^{M^e},\{U^e\}_{k,j=0}^{M^e}, 
%\{L^o_k\}_{k=0}^{M^o},\{U^o\}_{k,j=0}^{M^o}$
%\EndProcedure
%\end{algorithmic}
%\end{algorithm}
%
%\begin{algorithm}
%\caption{Solve of $H_{kj}u_j = b_k$, where $H$ has previously been LU 
%factorized for odd even coefficients using Algorithm (\ref{alg:lu}).}
%\label{alg:lusolve}
%\begin{algorithmic}[1]
%\Procedure{LUSolveoddeven}{}
%  \State \textbf{Input:} $\{U_{kj}\}_{k,j=0}^{M}, \{L_k\}_{k=0}^{M}, 
%\{b_k\}_{k=0}^{M}$ 
%  \State // Solve $Ly=b$ by forward elimination 
%  \State $y_0 \gets b_0$
%  \For{$k = 1$ \textbf{to} $M$}
%    \State $y_k \gets b_k - L_ky_{k-1}$
%  \EndFor
%  \State // Solve $Uu=y$ with back substitution
%  \State double $s = 0.0$     
%  \State $u_M = y_M / U_{MM}$    
%  \For{$i = M-1$ \textbf{to} $0$ \textbf{step} $-1$}
%     \State $u_i \gets y_i - U_{i, i+1}\cdot u_{i+1}$
%     \If{$i < M-1$}
%       \State $s \gets s + u_{i+2}$
%       \State $u_i \gets u_i - s\cdot U_{i,i+2}$            
%     \EndIf
%     \State $u_i \gets  u_i / U_{i,i}$
%   \EndFor
%  \State \Return $\{u_k\}_{k=0}^{M}$  
%\EndProcedure
%\end{algorithmic}
%
%\begin{algorithmic}[1]
%\Procedure{LUSolve}{}
%  \State \textbf{Input: $\{u_k\}_{k=0}^{N-2}, \{b_k\}_{k=0}^{N-2}, 
%\{U_{kj}^{e}\}_{k,j=0}^{M^{e}}, \{U_{kj}^{o}\}_{k,j=0}^{M^{o}}, 
%\{L^{e}_k\}_{k=0}^{M^{e}}, \{L^{o}_k\}_{k=0}^{M^{o}}$}
%  \State $\{u_{2k}\}_{k=0}^{M^e} \gets $ LUSolveoddeven($U^{e}, L^e, 
%\{b_{2k}\}_{k=0}^{M^e}$)
%  \State $\{u_{2k+1}\}_{k=0}^{M^o} \gets $ LUSolveoddeven($U^{o}, L^o, 
%\{b_{2k+1}\}_{k=0}^{M^o}$)
%  \State \Return $\{u_k\}_{k=0}^{N-2}$
%\EndProcedure
%\end{algorithmic}
%\end{algorithm}

The biharmonic matrix $\N{H}$ in Eq.~(\ref{eq:Biharmonic_matrix}) is more 
challenging to solve for efficiently, but 
it is still possible to find an algorithm that is $O(N)$ for a system of $N$ 
unknowns. Note 
that $\N{H}$ is the sum of three matrices $\N{Q}, \N{A}$ and $\N{B}$ and can be 
written as
\begin{equation}
\N{H} = c_0\N{Q} + c_1\N{A} + c_2 \N{B},
\end{equation}
where $c_0, c_1$ and $c_2$ are constants. The matrix $\N{A}_{kj}$ contains only 
three nonzero diagonals at $ j = k-2, k, k+2$,  whereas matrix $\N{B}_{kj}$ 
contains five 
nonzero diagonals at $ j = k-4, k-2, k, k+2, k+4$. The nonzero elements of the 
upper triangular $\N{Q}_{kj}$ matrix are given as
\begin{align}
 \N{Q}_{kk} &= -4(k+1)(k+2)^2, \\
 \N{Q}_{kj} &= p_kq_j + r_ks_j, \quad j = k+2, k+4, \ldots, N_x-4,
\end{align}
where 
\begin{align}
p_k &= 8 k (k+1)(k+2)(k+4)\pi, &q_j &= \frac{1}{j+3}, \label{eq:pk} \\
r_k &= 24(k+1)(k+2)\pi, &s_j &=  \frac{(j+2)^2}{(j+3)}. \label{eq:rk}
\end{align} 
For each row in $\N{H}_{kj}$ we have that $\N{H}_{kj}=c_0\N{Q}_{kj}\, \forall j 
> k+4$. A straight forward direct LU decomposition of $\N{H}$ can be performed 
as shown in Alg. 
\ref{alg:lu_biharmonic}, which leads to a lower triangular 
matrix $\N{L}_{kj}$ with two nonzero diagonals at $j=k-2$ and $k-4$ plus a 
unity 
main diagonal. The upper triangular matrix $\N{U}_{kj}$ is dense and as such a 
show stopper for an efficient back solve. However, we note that $\N{U}_{kj}$ 
contains only three distinct 
diagonals at $j=k, k+2$ and $k+4$, whereas the remaining part of $\N{U}_{kj}$ 
can be expressed as
\begin{equation}
\N{U}_{kj} = c_0(a_k q_j + b_k s_j), \quad j = k+6, k+8, \ldots, N_x-4, 
\label{eq:ab}
\end{equation}
where $\bm{a}$ and $\bm{b}$ are two new vectors that can be computed 
recursively from $\N{L}$ 
\begin{align}
a_k &= p_k, \, &\forall \, k= 0, 1, \notag \\
b_k &= r_k, \, &\forall \, k= 0, 1, \notag\\
a_k &= p_k - \N{L}_{k,k-2}a_{k-2}, \, &\forall \, k = 2, 3, \notag\\
b_k &= r_k - \N{L}_{k,k-2}b_{k-2}, \, &\forall \, k = 2, 3, \notag\\
a_k &= p_k - \N{L}_{k,k-2}a_{k-2} - \N{L}_{k,k-4}a_{k-4}, \, &\forall \, k = 4, 
5, 
\ldots, N, \notag\\
b_k &= r_k - \N{L}_{k,k-2}b_{k-2} - \N{L}_{k,k-4}b_{k-4}, \, &\forall \, k = 4, 
5, 
\ldots, N.
\end{align}
If $\bm{a}$ and $\bm{b}$ 
are pre-computed 
the total storage requirement for the complete LU decomposition is no more than 
$7N$, i.e., two diagonals for $\N{L}$, three diagonals for $\N{U}$ plus 
$\bm{a}$ and 
$\bm{b}$. The solution of the complete system can be obtained very quickly with 
one simple forward elimination and a back solve, where the back 
solve is very fast ($O(N)$) because of the formulation (\ref{eq:ab}), leading 
to a formula for the backwards substitution like (valid for $k \le N_x-8$)
\begin{equation}
u_k = \left(y_k - \N{U}_{k, k+2} u_{k+2} - \N{U}_{k, k+4} u_{k+4} - 
c_0 a_k\sum_{\substack{j=k+6\\k-j \text{ even}}}^{N_x-4} q_j u_j - 
c_0 b_k\sum_{\substack{j=k+6\\k-j \text{ even}}}^{N_x-4} 
s_j u_j\right)/\N{U}_{kk},
\end{equation}
that only requires 
one new addition to the row-sums for each back-traversed
row, see Alg. (\ref{alg:SolveBiharmonic}).

\begin{algorithm}
	\caption{LU decomposition of biharmonic operator $\N{H}$. Current algorithm 
	is using dense storage of $\N{U}_{kj}$, but no more than the 3 diagonals 
	need 
	to be stored eventually.}
	\label{alg:lu_biharmonic}
	\begin{algorithmic}[1]
		\Procedure{LUbiharmonic}{}
			\State \textbf{Input:} $\{\N{H}_{kj}\}_{k, j = 0}^{N}$
	        \State $\N{U}_{0j} = \N{H}_{0j} \quad \forall \, j \in 0, 2, 
	        \ldots, N$
	        \State $\N{U}_{1j} = \N{H}_{1j} \quad \forall \, j \in 1, 3, 
	        \ldots, N$
	        \State $\N{L}_{20} = \N{H}_{20}/\N{U}_{00}$
	        \State $\N{L}_{31} = \N{H}_{31}/\N{U}_{11}$
	        \State $\N{U}_{2j} = \N{H}_{2j} - \N{L}_{20} \N{U}_{0j} \quad 
	        \forall \, j \in 2, 4, \ldots, N$
	        \State $\N{U}_{3j} = \N{H}_{3j} - \N{L}_{31} \N{U}_{1j} \quad 
	        \forall \, j \in 
	        3, 5, \ldots, N$
	        \For{$k = 4, 5, \ldots, N$}
		        \State $\N{L}_{k,k-4} = \N{H}_{k, k-4} / \N{U}_{k-4, k-4}$
		        \State $\N{U}_{k,j} = \N{H}_{k,j} - \N{L}_{k, k-4} 
		        \N{U}_{k-4, j} 
		        \quad \forall j = k-2, k, \ldots, N$
	            \State $\N{L}_{k,k-2} = {\N{U}}_{k,k-2}/{\N{U}}_{k-2, k-2}$
		        \State $\N{U}_{kj} = {\N{U}}_{kj} - \N{L}_{k, k-2} \N{U}_{k-2, 
		        j} \quad 
		        \forall j = k, k+2, \ldots, N$
	        \EndFor
	        \State \Return $\{\N{L}_{kj}\}_{k, j = 0}^{N}$, $\{\N{U}_{kj}\}_{k, 
	        j = 
	        0}^{N}$
		\EndProcedure	
	\end{algorithmic}
\end{algorithm}

%\begin{algorithm}
%	\caption{Recursive formula to compute $\{a_k\}_{k=0}^{N_x-4}, 
%	\{b_k\}_{k=0}^{N_x-4}$ in Eq. (\ref{eq:ab}). Note that the lower triangular 
%	matrix $\{L_{kj}\}_{k,j=0}^{N_x-4}$ contains only two nonzero diagonals and 
%	the parameters $p_k$ and $r_k$ are given in Eqs. (\ref{eq:pk}) and 
%	(\ref{eq:rk}) respectively. }
%    \label{alg:ab}
%	\begin{algorithmic}[1]
%		\Procedure{ReduceAB}{}
%		\State \textbf{Input:} $ \{\N{L}_{kj}\}_{k,j=0}^{N_x-4}, c_0$
%		\For{$k=0,1$}
%		     \State $a_k \gets p_k$
%		     \State $b_k \gets r_k$
%		\EndFor
%		\For{$k=2,3$}
%		    \State $a_k \gets p_k - \N{L}_{k,k-2}\,a_{k-2}$
%		    \State $b_k \gets r_k - \N{L}_{k, k-2}\,b_{k-2}$
%		\EndFor
%		\For{$k=4, 5, \ldots, N_x-4$}
%     		\State $a_k \gets p_k - \N{L}_{k,k-2}\,a_{k-2} - \N{L}_{k, k-4} 
%\,a_{k-4}$
%	    	\State $b_k \gets r_k - \N{L}_{k, k-2}\,b_{k-2}- \N{L}_{k, k-4} 
%\,b_{k-4}$
%		\EndFor
%		\For{$k=0, 1, \ldots, N_x-4$}
%		    \State $a_k \gets a_k\cdot c_0$
%		    \State $b_k \gets b_k \cdot c_0$
%		\EndFor
%		\State \Return $\{a_k\}_{k=0}^{N_x-4}, \{b_k\}_{k=0}^{N_x-4}$
%        \EndProcedure
%\end{algorithmic}
%\end{algorithm}

\begin{algorithm}
	\caption{Solve biharmonic Eq. (\ref{eq:ufin}) with pre-computed $\N{L}$, 
	$\N{U}$ 
	matrices. The parameters $q_j$ and $s_j$ are given in Eqs. (\ref{eq:pk}) 
	and 
	(\ref{eq:rk}) respectively. }
	\label{alg:SolveBiharmonic}
	\begin{algorithmic}[1]
		\Procedure{LUSolveBiharmonic}{}
		  \State \textbf{Input:} $\{\N{U}_{kj}\}_{k,j=0}^{N}, 
		  \{\N{L}_{kj}\}_{k,j=0}^{N}, 
		  \{f_k\}_{k=0}^{N}, \{a_k\}_{k=0}^{N}, \{b_k\}_{k=0}^{N}$ 
		  \State // Solve $\N{L}_{kj}y_j=f_k$ by forward elimination 
		  \State $y_0 \gets f_0$
		  \State $y_1 \gets f_1$
		  \State $y_2 \gets f_2 - \N{L}_{20}y_{0}$
		  \State $y_3 \gets f_3 - \N{L}_{31}y_{1}$
		  \For{$k = 4$ \textbf{to} $N$}
		  \State $y_k \gets f_k - \N{L}_{k,k-2}y_{k-2} - \N{L}_{k, k-4}y_{k-4}$
		  \EndFor
		  \State // Solve $\N{U}_{kj}u_j=y_k$ with back 
		  substitution                    
		  \State $u_N = y_N / \N{U}_{NN}$
	      \State $u_{N-1} = y_{N-1} / \N{U}_{N-1,N-1}$
	      \State $u_{N-2} = \left(y_{N-2} - \N{U}_{N-2, 
	      N}y_N\right)/\N{U}_{N-2,N-2}$
	      \State $u_{N-3} = \left(y_{N-3} - \N{U}_{N-3, 
	      N-1}y_{N-1}\right)/\N{U}_{N-3,N-3}$

	      \State $q^o = q^e = s^o = s^e = 0$
		  \For{$k = N-4$ \textbf{to} $0$ \textbf{step} $-1$}
		  \State $u_k \gets y_k - \N{U}_{k, k+2} u_{k+2} - \N{U}_{k, k+4} 
		  u_{k+4} $ 
		  \If{$k < N-5$}
		  \State $j=k+6$
		  \If{$k$ is odd}
		    \State $q^o \gets q^o + u_{j}q_{j}$
		    \State $s^o \gets s^o + u_{j}s_{j}$
		    \State $u_k \gets u_k - q^o a_k - s^o b_k $	    
		  \Else
		    \State $q^e \gets q^e + u_{j}q_{j}$
		    \State $s^e \gets s^e + u_{j}s_{j}$	  
		    \State $u_k \gets u_k - q^e a_k - s^e b_k $
		  \EndIf
		  \EndIf
		  \State $u_k \gets  u_k / \N{U}_{kk}$
		  \EndFor
		  \State \Return $\{u_k\}_{k=0}^{N}$

		\EndProcedure
	\end{algorithmic}
\end{algorithm}

\section{Verification}
We have described a Navier-Stokes solver for channel flows that scales no worse 
than the fast transforms used to go between spectral and real space. All the 
matrices that make up the linear systems can be inverted using sparse direct LU 
decompositions that scales like $O(N)$. To verify and to further show that we 
can use these algorithms for large scale simulations, we will first study some 
simple 1D problems.

We first consider the roundoff error of the matrix decomposition methods 
described in Sec.~\ref{sec:implementation}. The roundoff error may be 
approximated 
by the following approach: Let $\{u_i\}_{i=0}^{N-4}$ be a uniformly distributed 
random vector. Compute the matrix vector product $\bm{f}=\N{H}\bm{u}$ and 
then solve $\N{H} \bm{v} = \bm{f}$, using Alg. \ref{alg:SolveBiharmonic}, for 
$\{v_i\}_{i=0}^{i=N-4}$. The roundoff error of the matrix decomposition 
algorithm may then be found as $\max_{0 \ge i \ge N-4}(|u_i-v_i|)/\max_{0 \ge i 
\ge N-4}|u_i|$. For this experiment we assume that $\triangle t=0.0001$ 
and $\nu=1/1200$, that are reasonable parameters for a turbulent channel flow 
simulation at $\text{Re}_{\tau}=1200$. Three different wavenumbers 
$z=(2, 200, 5600)$ are 
also chosen as representative for simulations that are using up to 8000 
wavenumbers in each periodic direction, which is approximately the resolution 
of the largest known channel simulations, performed by Lee and 
Moser~\cite{leemoser15}. The mesh in the wall normal direction 
is varied from 64 to 4096 points, which covers almost three times higher 
resolution than used by Lee and Moser \cite{leemoser15}. Results shown in the 
first three rows of Table \ref{tab:roundoff} indicate that 
roundoff errors are not very pronounced, even for problems as large as 
$N_x=4096$. Further, the roundoff errors increase only slowly with increasing 
the wavenumbers. Corresponding results are also computed for the Helmholtz 
matrix (\ref{eq:Helmholtz_matrix}) and 
results are shown in the last three rows of Table \ref{tab:roundoff}. For 
the Helmholtz matrix the roundoff errors are seen to be insignificant.

\begin{table}
	\caption{Roundoff error of the matrix decomposition method. The first three 
	rows show results for Alg. (\ref{alg:SolveBiharmonic}) and matrix $\N{H}(z, 
	\triangle t, \nu)$, whereas the last three rows show results for matrix 
	$\D{H}(z, \triangle t, \nu)$. Parameters 
	are chosen as $\nu=1/1200$ and $\triangle t=0.0001$ and the numbers shown 
	are for an average of 100 runs. }
	\label{tab:roundoff}
\begin{tabular}{cccccccc}
	   \multicolumn{8}{c}{$N_x$} \\
	   \hline
	$z$ & 64 & 128 & 256 & 512 & 1024 & 2048 & 4096 \\
	\hline
0 & 2.66e-15 & 6.71e-15 & 1.73e-14 & 4.64e-14 & 1.18e-13 & 3.30e-13 & 9.21e-13 
\\
200 & 8.39e-15 & 1.78e-14 & 5.12e-14 & 1.56e-13 & 4.47e-13 & 1.18e-12 & 
3.45e-12 \\
5600 & 7.09e-15 & 2.08e-14 & 5.65e-14 & 1.43e-13 & 4.38e-13 & 1.01e-12 & 
2.61e-12 \\
		\hline
0 & 3.29e-15 & 6.07e-15 & 6.05e-15 & 6.50e-15 & 5.63e-15 & 5.43e-15 & 5.13e-15 
\\
200 & 3.03e-15 & 6.30e-15 & 6.14e-15 & 5.87e-15 & 5.67e-15 & 5.50e-15 & 
5.39e-15 \\
5600 & 3.35e-15 & 7.78e-15 & 8.86e-15 & 7.94e-15 & 7.18e-15 & 7.20e-15 & 
7.03e-15

\end{tabular}
\end{table}

To show that the solvers are scaling with size, we investigate the execution time 
for solving $\N{H} \bm{v} = \bm{f}$ and $\D{H} \bm{v} = \bm{f}$ on a complete 
wavenumber mesh of size $(N_x,64,64)$. The numbers have been produced on a MacBook Pro laptop with the 2.8 GHz intel Core i7 processor. The average time for one single 
solve is shown in Table~\ref{tab:timings}, where the numbers in parenthesis are used to indicate scaling (unity would be perfectly $O(N)$). The scaling is seen to be close to linear, and variations 
are most likely due to how well datastructures fit in the cache. Naturally, the timings are not 100$\%$ reproducible, but the ability to solve large scale problems is evident.
\begin{table}
	\centering
	\caption{Execution time (in seconds) for one solve  of the Helmholtz 
	(\ref{eq:gfin}) and Biharmonic 
	(\ref{eq:ufin}) problems. The numbers in parenthesis are computed as $t_n / t_{n-1} / 2$, where $t_n$ is the current solvers timing in row $n$.}
	\label{tab:timings}
	\begin{tabular}{ccc}
%		$N_x$ & Biharmonic & Helmholtz \\
%		\hline
%64 & 6.49e-07 (*) & 4.64e-07 (*) \\ 
%128 & 1.40e-06 (1.08) & 9.43e-07 (1.02) \\ 
%256 & 2.96e-06 (1.06) & 1.93e-06 (1.02) \\ 
%512 & 6.09e-06 (1.03) & 4.01e-06 (1.04) \\ 
%1024 & 1.49e-05 (1.22) & 9.83e-06 (1.23) \\ 
%2048 & 2.53e-05 (0.85) & 1.64e-05 (0.83) \\ 
%4096 & 6.15e-05 (1.22) & 4.09e-05 (1.25) \\ 
%8192 & 1.25e-04 (1.02) & 8.18e-05 (1.00) 
Nx & Biharmonic & Helmholtz \\ 
\hline
64 & 8.09e-07 (0.00) & 4.51e-07 (0.00) \\ 
128 & 1.57e-06 (0.97) & 9.41e-07 (1.04) \\ 
256 & 2.76e-06 (0.88) & 1.97e-06 (1.05) \\ 
512 & 4.96e-06 (0.90) & 4.22e-06 (1.07) \\ 
1024 & 1.18e-05 (1.18) & 8.40e-06 (1.00) \\ 
2048 & 2.37e-05 (1.01) & 1.79e-05 (1.07) \\ 
4096 & 4.82e-05 (1.02) & 3.78e-05 (1.05) \\ 
8192 & 9.72e-05 (1.01) & 7.91e-05 (1.05)
	\end{tabular}
\end{table}

%\section{The incremental pressure correction scheme}
%
%The Navier Stokes equations are discretized in time using a second order 
%incremental pressure correction scheme with Crank-Nicolson on diffusion and 
%explicit Adams Bashforth on the nonlinear convection. The incremental pressure 
%correction scheme (IPCS) that is used consists of a tentative momentum 
%equation 
%followed by a pressure correction. The two steps may be performed iteratively. 
%The equations are iterated forward in time in equal size time steps $\triangle 
%t$ from initial conditions $u(\bm{x}, 0) = u^0(\bm{x})$ at $t=0$, such that 
%$t_n = n\triangle t$, $n \in \mathcal{Z}$. For turbulent channel flows initial 
%conditions are simply a perturbed state used to get a turbulent flow going. 
%
%The IPCS attempts to find velocity vector $\bm{u}^{n}$ on time step $t_n$  and 
%pressure midway between $t_n$ and $t_{n-1}$, $p^{n-1/2}$, given the solutions 
%on all previous time steps. The three steps that are required solved on each 
%time step are
%\begin{align}
%i)&\,\, \frac{\bm{u}^{*}-\bm{u}^{n-1}}{\triangle t} + \bm{N}^{n-1/2}   = \nu 
%\nabla^2 \bm{u}^{n-1/2} - \nabla{p}^{n-3/2} + \bm{f}, \quad \bm{u}^*=0 \text{ 
%for } x=\pm1   \label{eq:NSCN} \\
%ii)&\,\,\begin{cases}
%  \frac{\bm{u}^{n}-\bm{u}^*}{\triangle t} &= -\nabla p^* \\
%  \nabla \cdot \bm{u}^{n} &= 0, \quad \bm{u}^n\cdot \bm{n} = 0,\text{ for } 
%x=\pm1
%  \end{cases} \label{eq:momupdate}\\
%  iii)&\,\,p^{n-1/2}=p^{n-3/2}+p^* \label{eq:pupdate} 
%\end{align}
%Here $\bm{u}^{n-1/2} = 0.5(\bm{u}^*+\bm{u}^{n-1})$, $\bm{u}^*$ is a tentative 
%velocity, $p^*$ is a pressure correction, and the explicit convection is 
%computed as $\bm{N}^{n-1/2}=1.5(\bm{u}^{n-1}\cdot \nabla) \bm{u}^{n-1} - 
%0.5(\bm{u}^{n-2}\cdot \nabla) \bm{u}^{n-2}$. The second step is computed by 
%solving a Poisson equation for the pressure correction
%\begin{equation}
%  \nabla^2p^* = \frac{\nabla \cdot \bm{u}^*}{\triangle t}
%\end{equation}
%and subsequently updating the velocity through (\ref{eq:momupdate}).
%
%All three velocity components use the Dirichlet basis (\ref{eq:u_sol}), 
%whereas the pressure is searched for in $\N{V}_N$
%\begin{equation}
%p(\bm{x}, t) = \frac{1}{N_yN_z}\sum_{\bm{k} \in \N{K}_N} \hat{p}_{\bm{k}}(t) 
%\N{\psi}_{\bm{k}}(\bm{x}), \label{eq:p_sol}
%\end{equation}
% 
%The spectral Galerkin method becomes: 
%
%1) Find $\bm{u}^*$ in $\D{V}_N^3$ by solving
%\begin{multline}
% \frac{1}{\triangle t}\left<\bm{u}^*-\bm{u}^{n-1}, \bm{\D{\psi}}\right>_w + 
%\left<\bm{N}^{n-1/2}, \bm{\D{\psi}} \right>_w   = \nu \left<\nabla^2 
%\bm{u}^{n-1/2}, \bm{\D{\psi}} \right>_w \\ - \left<\nabla{p}^{n-3/2}, 
%\bm{\D{\psi}} \right>_w + \left<\bm{f}, \bm{\D{\psi}} \right>_w \,\,  \forall 
%\, \bm{\D{\psi}} \in \D{V}_N^3. \label{eq:tentative}
%\end{multline}
%
%2) Compute the pressure correction $p^*\in \N{V}_N$ by solving
%\begin{equation}
%\left< \nabla^2 p^*, \N{\psi} \right>_w = \frac{1}{\triangle t} \left<\nabla 
%\cdot \bm{u}^*, \N{\psi} \right>_w\,\, \forall \, \N{\psi} \in \N{V}_N. 
%\label{eq:pcorr}
%\end{equation}
%
%3) Update the velocity, $\bm{u}^n \in \D{V}_N^3$, and pressure, $p^{n-1/2} \in 
%\N{V}_N$, by solving
%\begin{align}
%\left< \bm{u}^n, \bm{\D{\psi}}\right>_w &= \left< \bm{u}^*, 
%\bm{\D{\psi}}\right>_w - \triangle t \left< \nabla p^*, \bm{\D{\psi}} 
%\right>_w 
%\,\,  &\forall \, \bm{\D{\psi}} \in \D{V}_N^3, \label{eq:u_update} \\
%\left<p^{n-1/2}, \N{\psi}\right>_w &= \left<p^{n-3/2}, \N{\psi}\right>_w + 
%\left<p^*, \N{\psi}\right>_w \,\, &\forall \, \N{\psi} \in 
%\N{V}_N,\label{eq:pres_update}
%\end{align}
%where the pressure update simply resorts to $\hat{p}^{n-1/2} = \hat{p}^{n-3/2} 
%+ \hat{p}^* \, \forall \, \bm{k}\in \N{K}_N$.
%
%\section{Implementation}
%The vector valued equations (\ref{eq:tentative}) and (\ref{eq:u_update}) are 
%solved in a segregated manner, one component at the time. We use $u_i$ and 
%$\hat{u}_i$ for component $i$ of the real and transformed velocity vector. The 
%inner products in Eqs. (\ref{eq:tentative}, \ref{eq:pcorr}, \ref{eq:u_update}, 
%\ref{eq:pres_update}) lead to a range of new matrices
%\begin{align}
% \left< u_i^* - u_i^{n-1}, \D{\psi}\right>_w &= 
%h\D{B}\left({\hat{u}}_i^*-{\hat{u}}_i^{n-1} \right) \quad &i \in (0,1,2), \\ 
% \left<\nabla^2 {u}_i^{n-1/2}, \D{\psi}\right>_w &= -h  \left( \D{A} 
%+(\underline{m}^2+\underline{n}^2)\D{B}\right) \hat{u}_i^{n-1/2} \quad &i \in 
%(0,1,2), \\ 
% \left< f_i, \D{\psi}\right>_w &= h\D{\mathcal{S}}(f_i) &i \in (0,1,2), \\
% \left<N_i^{n-1/2}, \D{\psi} \right>_w &= h\D{\mathcal{S}}(N_i^{n-1/2}) &i \in 
%(0,1,2), \\  
% \left< \nabla p^{n-3/2}, \bm{\D{\psi}} \right>_w &= h\left[\grave{C},\, 
%\imath \underline{m} \grave{B},\, \imath \underline{n} \grave{B} \right] 
%\hat{p}^{n-3/2}, \\ 
% \left< \nabla^2 p^*, \N{\psi} \right>_w &= -h\left(\N{A} +(\underline{m}^2 + 
%\underline{n}^2)\N{B} \right) \hat{p}^*, \\ 
% \left<\nabla \cdot \bm{u}^*, \N{\psi} \right>_w &= h\left( \tilde{C} 
%\hat{u}^* + \imath \underline{m} \tilde{B} \hat{v}^* + \imath \underline{n} 
%\tilde{B} \hat{w}^* \right), 
%\end{align}
%where $\D{A}_{kj} = -\left(\D{\phi}^{''}_{j}, \D{\phi}_{k} \right)_w, 
%\N{A}_{kj} = -\left(\N{\phi}^{''}_{j}, \N{\phi}_{k} \right)_w,  \grave{C}_{kj} 
%= \left(\N{\phi}^{'}_j, \D{\phi}_k \right)_w, \tilde{C}_{kj} = 
%(\D{\phi}^{'}_j, 
%\N{\phi}_k)_w, \N{B}_{kj} = \left( \N{\phi}_j, \N{\phi}_k \right)_w, 
%\grave{B}_{kj} = (\N{\phi}_j, \D{\phi}_k)_w$ and $\tilde{B}_{kj} = 
%(\D{\phi}_j, 
%\N{\phi}_k)_w$ are all more or less sparse matrices. Arranging the governing 
%equations on matrix form we obtain for the three tentative velocity components 
%in (\ref{eq:tentative})
%\begin{align}
%\D{H} \hat{u}^* &= \left(\frac{4}{\nu \triangle t} \D{B}-\D{H}\right) 
%\hat{u}^{n-1} - \left[\D{\mathcal{S}}(N_x^{n-1/2}) + \grave{C} \hat{p}^{n-3/2} 
%+ \D{\mathcal{S}}(f_x)\right]\frac{2}{\nu}, \label{eq:tentativeUA}\\
%\D{H} \hat{v}^* &=  \left(\frac{4}{\nu \triangle t} 
%\D{B}-\D{H}\right)\hat{v}^{n-1} - \left[\D{\mathcal{S}}(N_y^{n-1/2}) + \imath 
%\underline{m} \grave{B} \hat{p}^{n-3/2} + 
%\D{\mathcal{S}}(f_y)\right]\frac{2}{\nu}, \label{eq:tentativeVA}\\
%\D{H} \hat{w}^* &= \left(\frac{4}{\nu \triangle t} \D{B}-\D{H}\right) 
%\hat{w}^{n-1} - \left[\D{\mathcal{S}}(N_z^{n-1/2}) + \imath \underline{n} 
%\grave{B} \hat{p}^{n-3/2} + 
%\D{\mathcal{S}}(f_z)\right]\frac{2}{\nu},\label{eq:tentativeWA}
%\end{align}
%where $\D{H}=\D{A} +(\underline{m}^2 + \underline{n}^2 + \frac{2}{\nu 
%\triangle t})\D{B}$ is a matrix consisting of one subdiagonal, and where every 
%second diagonal is zero. As such, we may split the matrix into two smaller 
%(and 
%decoupled) matrices for odd and even coefficients $H^e_{k,j} = H_{2k,2j}$ and 
%$H^o_{k,j} = H_{2k+1,2j+1}$, where the two matrices $H^{e,o}$ are of type 
%upper 
%Hessenberg, with only 4 distinct values on each row. This structure of $H$ 
%allows us to perform a very efficient tailored LU decomposition, as shown in 
%Algorithm (\ref{alg:lu}), and a very efficient tailored solve through 
%Algorithm 
%(\ref{alg:lusolve}). Regarding Algorithm (\ref{alg:lusolve}), we note that the 
%ability to decouple the system into odd and even coefficients leads to two 
%subsystem that may be trivially solved simultaneously in two threads by 
%letting 
%one thread call LUODDEVEN with even coefficients, and the second thread call 
%it 
%for odd. 
%
%The linear system arising for the pressure correction equation is similar to 
%the tentative velocities
%\begin{equation}
%\N{H}\hat{p}^* = -\frac{1}{\triangle t}\left( \tilde{C} \hat{u}^* + \imath 
%\underline{m} \tilde{B} \hat{v}^* + \imath \underline{n} \tilde{B} \hat{w}^* 
%\right), \label{eq:pressureA}
%\end{equation}
%where $\N{H}=\N{A} +(\underline{m}^2 + \underline{n}^2)\N{B}$ is of the same 
%structure as $\D{H}$ and as such  should be factorizable in the same way. 
%However, entries of the matrix $\N{H}_{kj}$ contain the column index $j^2$ 
%(see 
%$\N{A}$), and as such the values on each row of the matrix $\N{H}_{kj}$ are 
%not 
%constant for $j \ge k+4$ and the LU decomposition of $\N{H}$ will require a 
%full upper $U$ matrix. To circumvent this we define $\N{H}_{kj} = 
%\underline{H}_{kl}J_{lj}$, where $J_{kj}= j^2\delta_{kj}$ (no summation 
%implied). We then set $w_k=J_{kj}\hat{p}_j$, use $b_k$ for the right hand side 
%of (\ref{eq:pressureA}) and solve
%\begin{align}
%  \underline{H}_{kj}w_j &= b_k &\forall \, k \in 1, 2, \ldots, N_x-2, \\
%  \hat{p}_k &= J_{kj}^{-1}w_j &\forall \, k \in 1, 2, \ldots, N_x-2.
%\end{align}
%Here the first system $\underline{H}_{kj} w_j = b_k$ can be solved using the 
%same LU factorization as the $\D{H}$ matrix, since 
%$\underline{H}_{kj}=\underline{H}_{k, k+4}\, \forall \, j=k+6, k+8, \ldots$ 
%and 
%there are only 4 distinct values on each row. The implementation is the same 
%as 
%for $\D{H}$, except that for the matrix $\underline{H}$ the indices start at 1 
%instead of zero.
%
%The velocity update is computed on matrix form as
%\begin{align}
%  \hat{u}^n &= \hat{u}^* - \triangle t \D{B}^{-1} \left( \grave{C} \hat{p}^* 
%\right),\\
%  \hat{v}^n &= \hat{v}^* - \triangle t \D{B}^{-1} \left( \imath \underline{m} 
%\grave{B} \hat{p}^* \right),\\ 
%  \hat{w}^n &= \hat{w}^* - \triangle t \D{B}^{-1} \left( \imath \underline{n} 
%\grave{B} \hat{p}^* \right).
%\end{align}
%using a tridiagonal solver.
%
%The convection term is computed in standard form with a pseudo-spectral method 
%that requires a fast scalar product
%\begin{equation}
%  \D{\mathcal{S}}(N_i^{n-1}) = \D{\mathcal{S}}(\bm{u}^{n-1} \cdot \nabla 
%u_i^{n-1}) \quad \forall \, i \in (0, 1, 2).
%\end{equation}
%We obtain the gradients of $u_i$ in real space by projecting $\partial 
%u/\partial x$ to $\D{V}_{N_x}$ (boundary condition $\partial u/\partial x(\pm 
%1)=0$ follows from continuity) and $\partial v/\partial x$ and $\partial 
%w/\partial x$ to $V_{N_x}$. 
%\begin{align}
%  \frac{\partial u}{\partial x} &= \D{\mathcal{T}}^{-1} 
%(\D{B}^{-1}(\D{C}\hat{u})), \\
%  \frac{\partial v}{\partial x} &= \mathcal{T}^{-1} 
%(B^{-1}(\acute{C}\hat{v})), \\
%  \frac{\partial w}{\partial x} &= \mathcal{T}^{-1} 
%(B^{-1}(\acute{C}\hat{w})),  
%\end{align}
%where $\D{C}_{kj} = (\D{\phi}_j^{'}, \D{\phi}_k)_w$,  $\acute{C_{kj}} = 
%(\D{\phi}_j^{'}, T_k)_w$ and $B=(T_j, T_k)_w$. Partial derivatives with 
%respect 
%to periodic directions are computed as
%\begin{align}
% \frac{\partial u_i}{\partial y} &= \D{\mathcal{T}}^{-1}(\imath \underline{m} 
%\hat{u}_i) &\forall \, i \in (0, 1, 2), \\
% \frac{\partial u_i}{\partial w} &= \D{\mathcal{T}}^{-1}(\imath \underline{n} 
%\hat{u}_i) &\forall \, i \in (0, 1, 2).
%\end{align}
%Dealiasing is implemented using the 2/3-rule.
%
%We finally note that the right hand sides of all the linear systems that need 
%to solved can be assembled fast using tailored matrix vector products taking 
%advantage of the fact that there are at most three distinct diagonals in any 
%matrix. This concludes the implementation required to solve the Navier-Stokes 
%equations by the incremental pressure correction method and the spectral 
%Shen-Fourier Galerkin method. We note that all terms in assembly and solve may 
%be performed using fast routines, at worst scaling like $O(N_x \log_2 N_x)$. 
%
%\section{Serial scaling}
%
%\section{Parallel scaling}
%
%
%

\bibliography{bib.bib}

\section{Appendices}
%\subsection*{Fast Shen transforms}
%The fast forward and inverse Chebyshev transforms are used by the Shen 
%transforms. The code below computes both scalar products and complete 
%transforms for both 
%
%\begin{python}
%from numpy import *
%from scipy.fftpack import dct
%
%def fct(f, transform="GL"):
%    """Fast Chebyshev transform."""
%    N = f.shape[0]
%    f_hat = zeros(N)
%    if transform == "GL":
%        fk[:] = dct(fj, type=2, axis=0)            
%        fk /= N
%        fk[0] /= 2
%                
%    elif transform == "GC":
%        fk[:] = dct(fj, type=1, axis=0)/(N-1)
%        fk[0] /= 2
%        fk[-1] /= 2
%            
%    return fk
%
%def ifct(fk, transform="GL"):
%    """Inverse fast Chebyshev transform."""
%    if transform == "GL":
%        f = 0.5*dct(fk, type=3, axis=0)
%        f += 0.5*fk[0]
%    
%    elif transform == "GC":
%        f = 0.5*dct(fk, type=1, axis=0)
%        f += 0.5*fk[0]
%        f[::2] += 0.5*fk[-1]
%        f[1::2] -= 0.5*fk[-1]
%
%    return f
%
%def fastChebScalar(f, transform="GL"):
%    """Fast Chebyshev scalar product."""
%    N = f.shape[0]
%    if transform == "GL":
%        fk = dct(fj, type=2, axis=0)*pi/(2*N)
%        
%    elif transform == "GC":
%        fk = dct(fj, type=1, axis=0)*pi/(2*(N-1))
%    return fk
%
%def fastShenScalar(f, transform="GL"):
%    """Fast Shen scalar product"""                
%    fk = fastChebScalar(f, transform)
%    fk[:-2] -= fk[2:]
%    return fk
%
%def ifst(fk):
%    """Fast inverse Shen transform
%    """
%    N = len(fj)
%    wk = np.zeros(N, dtype=fk.dtype)        
%    wk[:-2] = fk[:-2] 
%    wk[2:] -= fk[:-2] 
%    return ifct(wk)
%    
%
%def fst(fj, transform="GL"):
%    """Fast Shen transform"""
%    fk = fastShenScalar(fj, transform)
%    
%    N = fj.shape[0]
%    if transform == "GL":
%        ck = np.ones(N-2); ck[0] = 2
%        
%    elif transform == "GC":
%        ck = np.ones(N-2); ck[0] = 2; ck[-1] = 2
%        
%    a = np.ones(N-4)*(-np.pi/2)
%    b = np.pi/2*(ck+1)
%    c = a.copy()
%    fk[:-2] = TDMA_1D(a, b, c, fk[:-2])
%        
%    return fk
%
%def fastShenScalar(f, transform="GL"):
%    """Fast Shen scalar product Neumann basis"""        
%    k  = wavenumbers(f.shape)
%    fk = fastChebScalar(f, transform)
%    fk[:-2] -= ((k/(k+2))**2) * fk[2:]
%    return fk
%
%def ifst(fk):
%    """Fast inverse Shen scalar transform Neumann basis"""
%    k = self.wavenumbers(fk.shape[0])
%    wk = np.zeros(fk.shape[0])    
%    wk[1:-2] = fk[1:-2]
%    wk[3:] -= (k[1:]/(k[1:]+2))**2*fk[1:-2]
%    f = ifct(wk, transform)
%    return f
%        
%def fst(f, transform="GL"):
%    """Fast Shen transform Neumann basis"""
%    fk = fastShenScalar(f, transform)
%    N = fj.shape[0]
%    k = wavenumbers(N)
%    ck = np.ones(N-3)
%    if transform == "GC": ck[-1] = 2
%    a = np.ones(N-5)*(-np.pi/2)*(k[1:-2]/(k[1:-2]+2))**2
%    b = np.pi/2*(1+ck*(k[1:]/(k[1:]+2))**4)
%    c = a.copy()
%    fk[1:-2] = TDMA_1D(a, b, c, fk[1:-2])
%    return fk
%
%# Cython implementation of TDMA    
%import numpy as np
%cimport cython
%cimport numpy as np
%#cython: boundscheck=False
%#cython: wraparound=False
%
%ctypedef fused T:
%    np.float64_t
%    np.complex128_t
%
%def TDMA_1D(np.ndarray[np.float64_t, ndim=1] a, 
%            np.ndarray[np.float64_t, ndim=1] b, 
%            np.ndarray[np.float64_t, ndim=1] c, 
%            np.ndarray[T, ndim=1] d):
%    cdef:
%        unsigned int n = b.shape[0]
%        unsigned int m = a.shape[0]
%        unsigned int k = n - m
%        int i
%        
%    for i in range(m):
%        d[i + k] -= d[i] * a[i] / b[i]
%        b[i + k] -= c[i] * a[i] / b[i]
%    for i in range(m - 1, -1, -1):
%        d[i] -= d[i + k] * c[i] / b[i + k]
%    for i in range(n):
%        d[i] /= b[i]
%        
%    return d
%    
%
%\end{python}
\subsection*{Matrices}

\begin{equation}
 \{\D{A}_{kj}\}_{k,j=0}^{N_x-2} = -\left(\D{\phi}^{''}_{j}, \D{\phi}_{k} \right)_w = \begin{cases}
 2\pi(k+1)(k+2), &j=k,\\
 4\pi(k+1), & j=k+2, k+4, k+6, \ldots, \\
 0, &\text{otherwise}.
 \end{cases}
\end{equation}

\begin{equation}
 \{\N{A}_{kj}\}_{k,j=1}^{N_x-2} = -\left(\N{\phi}^{''}_{j}, \N{\phi}_{k} \right)_w = \begin{cases}
 2\pi j^2(k+1)/(k+2), &j=k,\\
 4\pi j^2(k+1)/(k+2)^2, & j=k+2, k+4, k+6, \ldots, \\
 0, &j>k, \text{ or } k+j \text{ odd}.
 \end{cases}
\end{equation}

\begin{equation}
 \{\N{C}_{kj}\}_{k,j=0}^{N_x-4, N_x-2} = \left(\D{\phi}^{'}_j, \N{\phi}_k 
 \right)_w = \begin{cases}
 -\pi(k+1), &j=k-1,\\
 2\pi(k+1), & j=k+1, \\
 -\pi(k+1), & j=k+3, \\
 0, &\text{otherwise}.
 \end{cases}
\end{equation}

\begin{equation}
 \{\D{C}_{kj}\}_{k,j=0}^{N_x-2, N_x-4} = \left(\N{\phi}^{'}_j, \D{\phi}_k 
 \right)_w = \begin{cases}
 \pi (k-2)(k+1)/k, &j=k-3,\\
 -2 \pi(k+1)^2/(k+2), & j=k-1, \\
 \pi(k+1), & j=k+1, \\
 0, &\text{otherwise}.
 \end{cases}
\end{equation}

\begin{equation}
 \{\N{B}_{kj}\}_{k,j=0}^{N_x-4} = \{\N{B}_{jk}\}_{j,k=0}^{N_x-4} = (\N{\phi}_j, 
 \N{\phi}_k)_w = \begin{cases}
 \frac{\pi}{2} \left(c_k + 4 \left(\frac{k+2}{k+3} \right)^2 + c_{k+4} 
 \left(\frac{k+1}{k+3}\right)^2    \right), &j=k,\\
 -\pi \left( \frac{k+2}{k+3} + \frac{(k+4)(k+1)}{(k+5)(k+3)} \right), &j=k+2,\\
 \frac{\pi}{2} \frac{k+1}{k+3} , & j=k+4, \\
 0, &\text{otherwise}.
 \end{cases}
\end{equation}

\begin{equation}
 \{B_{kj}\}_{k,j=0}^{N_x} = (T_k, T_j)_w = \frac{c_k \pi}{2} \delta_{kj}
\end{equation}

For Gauss-Lobatto points we have $c_0=c_{N_x}=2$ and $c_k=1$ for $0<k<N_x$. For Gauss-Chebyshev points we have $c_0=2$ and $c_k=1$ for $k>0$. Note that for an $L^2_w$ scalar product we would have $c_{N_x}=1$ as well, but for the $l^2_w$ scalar product used here we get $c_{N_x}=2$. This disagreement, that follows from inexact quadrature at the highest mode, is not reported by Shen, but is reported by, e.g., Kopriva.

\end{document}